---
title: "Punishment Study 2.1.2 — Main Analysis"
subtitle: "Confirmatory and Exploratory Tests of the Punitiveness Ecosystem"
author: "David G. Kamper"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    number_sections: false
  pdf_document:
    toc: true
---

```{r knitr_setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 7
)
```

# Overview

This script performs the **complete main analysis** for the Punishment Study 2.1.2,
covering data cleaning through hypothesis testing through exploratory follow-ups.

## Pre-registered hypotheses

  - **H1**: Punitiveness is positively correlated with all correlate measures.
  - **H2**: The hostile aggression cluster correlates more strongly with
    punitiveness than the crime concerns cluster.
  - **H3**: Most correlate measures are positively intercorrelated, indicating
    a cohesive "punitiveness mindset."

## Phases

  1. **Phase 0** — Setup and package loading
  2. **Phase 1** — Data loading, filtering, missing-data check
  3. **Phase 2** — Reverse coding (metadata-driven)
  4. **Phase 3** — Z-score sentencing decisions within vignette
  5. **Phase 4** — Composite construction with Cronbach's α
  6. **Phase 5** — Demographics and descriptive statistics
  7. **Phase 6** — Confirmatory hypothesis tests (H1, H2, H3) + FDR correction
  8. **Phase 7** — Correlation tables for export
  9. **Phase 8** — Vignette stability sensitivity analyses
  10. **Phase 9** — Exploratory analyses (process violations, blood sports,
      revenge vs. vengefulness, EFA, CFA, demographic moderators)
  11. **Phase 10** — Multiple regression (clusters, constructs, hierarchical)
  12. **Phase 11** — Save all output files

## Data note

This script operates on `Data_Final.csv`, which contains 95 variables across
~625 raw responses. The Qualtrics survey export (`PUNISHMENT_21_-__Pro-Social.qsf`)
references additional items (`revenge_4`, `proportionality`, `severity_1/2`,
`essential_1-4`, `choice_1/2`) that **are not present in `Data_Final.csv`**.
If these items are needed in the future, they must be sourced from a more
complete data export.

## Output organization

All outputs are written to subdirectories of the working directory:

  - `output/tables/` — CSV tables (correlations, descriptives, regression results)
  - `output/figures/` — PNG figures (heatmaps, histograms, scree plots)
  - `data/cleaned/` — Cleaned dataset for downstream use (e.g., NLP pipeline)

# PHASE 0: SETUP AND PACKAGE LOADING

```{r phase0_paths}
# === PATH CONFIGURATION ===
# All paths are relative to the working directory of this Rmd file.
# Set your working directory to the Analysis folder before knitting.

# Input data path
DATA_PATH <- "data/raw/Data_Final.csv"

# Output directories
OUT_TABLES  <- "output/tables"
OUT_FIGURES <- "output/figures"
OUT_DATA    <- "data/cleaned"

# Create output directories if they don't exist
for (dir in c(OUT_TABLES, OUT_FIGURES, OUT_DATA)) {
  if (!dir.exists(dir)) dir.create(dir, recursive = TRUE)
}

cat("Output directories ready:\n")
cat("  Tables:  ", normalizePath(OUT_TABLES), "\n")
cat("  Figures: ", normalizePath(OUT_FIGURES), "\n")
cat("  Cleaned: ", normalizePath(OUT_DATA), "\n")
```

```{r phase0_packages}
# === PACKAGE LOADING ===
# Each package's purpose is annotated for transparency.

required_packages <- c(
  "tidyverse",      # Data manipulation, visualization (dplyr, ggplot2, etc.)
  "psych",          # Cronbach's alpha, EFA, KMO, parallel analysis
  "corrplot",       # Correlation heatmaps
  "Hmisc",          # rcorr() for correlation matrices with p-values
  "cocor",          # Steiger's Z-test for comparing dependent correlations
  "lme4",           # Mixed-effects models (random intercepts)
  "lmerTest",       # P-values for mixed-effects models
  "car",            # Type-III ANOVA, regression diagnostics
  "effectsize",     # Effect size calculations (eta-squared, Cohen's d)
  "RColorBrewer",   # Color palettes for figures
  "knitr",          # Tables for knitted output
  "broom",          # tidy() for model outputs
  "scales",         # Percentage formatting
  "performance",    # ICC computation
  "lavaan"          # Confirmatory factor analysis
)

# Install missing packages
new_packages <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
if (length(new_packages)) install.packages(new_packages)

# Load all packages
invisible(lapply(required_packages, library, character.only = TRUE))

# Set numeric display options
options(scipen = 999)  # Avoid scientific notation in printed output
options(digits = 4)    # Default decimal places
```

```{r phase0_helpers}
# === HELPER FUNCTIONS ===
# Defined once here, used throughout the script.

#' Format a p-value for display
#'
#' @param p A numeric p-value
#' @return A character string
format_p <- function(p) {
  if (is.na(p)) return("NA")
  if (p < .001) return("< .001")
  return(sprintf("= %.3f", p))
}

#' Convert p-value to significance stars
#'
#' @param p A numeric p-value or vector
#' @return A character vector of stars
sig_stars <- function(p) {
  case_when(
    is.na(p) ~ "",
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ ""
  )
}

#' Compute composite score and Cronbach's alpha
#'
#' Wraps psych::alpha() for consistent reporting. Returns the composite
#' (row-mean of items) along with the reliability estimate.
#'
#' @param data Data frame containing the items
#' @param items Character vector of item column names
#' @param composite_name Display name for printout
#' @param print_output Whether to print formatted output
#' @return A list with composite, alpha, and n_items
compute_composite <- function(data, items, composite_name, print_output = TRUE) {

  # Select only the items of interest
  item_data <- data %>% select(all_of(items))

  # Compute Cronbach's alpha
  # check.keys = FALSE because we've already reverse-coded items manually
  alpha_result <- psych::alpha(item_data, check.keys = FALSE)
  alpha_value  <- alpha_result$total$raw_alpha

  # Compute composite as the mean of items (handles missing values pairwise)
  composite <- rowMeans(item_data, na.rm = TRUE)

  if (print_output) {
    cat(sprintf("  %-32s α = %.3f (k = %d)\n",
                composite_name, alpha_value, length(items)))
  }

  list(composite = composite,
       alpha     = alpha_value,
       n_items   = length(items))
}

#' Get correlation with 95% CI
#'
#' @param x Variable name (string)
#' @param y Variable name (string)
#' @param data Data frame
#' @return Named numeric vector
get_cor_ci <- function(x, y, data) {
  test <- cor.test(data[[x]], data[[y]])
  c(r        = unname(test$estimate),
    ci_lower = test$conf.int[1],
    ci_upper = test$conf.int[2],
    p        = test$p.value)
}

#' Format a correlation matrix with significance stars
#'
#' @param cor_matrix Matrix of correlations
#' @param p_matrix Matrix of p-values (same dims)
#' @param decimals Digits after decimal
#' @return Character matrix with stars appended
format_cor_table <- function(cor_matrix, p_matrix, decimals = 2) {
  # Flatten matrix to vector for sig_stars (case_when expects vectors,
  # not matrices), then reshape back to matrix
  r_formatted <- round(cor_matrix, decimals)
  stars_vec   <- sig_stars(as.vector(p_matrix))
  values_vec  <- paste0(format(as.vector(r_formatted), nsmall = decimals),
                        stars_vec)
  result      <- matrix(values_vec,
                        nrow = nrow(cor_matrix),
                        dimnames = dimnames(cor_matrix))
  diag(result) <- "-"
  result
}

cat("Helper functions loaded:\n")
cat("  format_p()         — p-value formatting\n")
cat("  sig_stars()        — p → stars\n")
cat("  compute_composite() — composite + alpha\n")
cat("  get_cor_ci()       — correlation with 95% CI\n")
cat("  format_cor_table() — formatted correlation matrix\n")
```


# PHASE 1: DATA LOADING AND CLEANING

## 1.1 Load raw data

```{r phase1_load}
# Read the authoritative data file. Data_Final.csv contains 95 columns,
# matching the variables collected and analyzed for the paper.
raw_data <- read.csv(DATA_PATH, stringsAsFactors = FALSE)

cat(sprintf("Raw data loaded: %d rows × %d columns\n", 
            nrow(raw_data), ncol(raw_data)))
cat(sprintf("File: %s\n", DATA_PATH))
```

## 1.2 Filter participants

We apply four pre-registered filters:

  1. `Progress == 100` (completed the survey)
  2. `Finished == 1` (Qualtrics-flagged completion)
  3. `AttnCheck1 == 9` ("Strongly agree" attention check)
  4. `AttnCheck2 == 4` (8 + 7 = 15, which is option 4 on the scale)

Each filter's exclusion count is reported for transparency.

```{r phase1_filter}
# Track sample size at each step for reporting
n_raw <- nrow(raw_data)

cat("=== DATA FILTERING ===\n\n")
cat(sprintf("Starting N: %d\n", n_raw))

# --- Filter 1: Survey completion (Progress) ---
df <- raw_data %>% filter(Progress == 100)
n_after_progress <- nrow(df)
cat(sprintf("After Progress = 100:  %d  (excluded: %d)\n",
            n_after_progress, n_raw - n_after_progress))

# --- Filter 2: Survey completion (Finished flag) ---
df <- df %>% filter(Finished == 1)
n_after_finished <- nrow(df)
cat(sprintf("After Finished = 1:    %d  (excluded: %d)\n",
            n_after_finished, n_after_progress - n_after_finished))

# --- Filter 3: Attention check 1 (instruction-following) ---
df <- df %>% filter(AttnCheck1 == 9)
n_after_attn1 <- nrow(df)
cat(sprintf("After AttnCheck1 = 9:  %d  (excluded: %d)\n",
            n_after_attn1, n_after_finished - n_after_attn1))

# --- Filter 4: Attention check 2 (basic arithmetic) ---
df <- df %>% filter(AttnCheck2 == 4)
n_after_attn2 <- nrow(df)
cat(sprintf("After AttnCheck2 = 4:  %d  (excluded: %d)\n",
            n_after_attn2, n_after_attn1 - n_after_attn2))

# --- Final sample ---
cat(sprintf("\n*** FINAL SAMPLE: N = %d ***\n", nrow(df)))
cat(sprintf("Total excluded: %d (%.1f%%)\n\n",
            n_raw - nrow(df), 100 * (n_raw - nrow(df)) / n_raw))
```

## 1.3 Missing data check

```{r phase1_missing}
# Spot-check missingness on key variables. Any non-zero counts here would
# indicate something unexpected upstream — these items are the survey backbone.
key_vars <- c("Sentence_1", "punishmore_1", "punishmore_2", "vignette",
              "hate_1", "revenge_1", "rwa_1", "sdo_1")

missing_summary <- df %>%
  select(all_of(key_vars)) %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "N_Missing")

cat("=== MISSING DATA CHECK (key variables) ===\n\n")
print(as.data.frame(missing_summary), row.names = FALSE)
cat("\n")
```

# PHASE 2: REVERSE CODING

Items are reverse-coded so that **higher values always indicate a more punitive
or more hostile response**. This makes downstream interpretation consistent.

The reverse-coding logic is captured in a single metadata table (`reverse_spec`)
which serves as the auditable source of truth for every reversed item.

```{r phase2_metadata}
# === REVERSE-CODING METADATA ===
# Each row documents one reverse-coded item:
#   - item:       column name in the data
#   - scale_max:  Likert scale ceiling (used to compute reversal: max+1 − x)
#   - construct:  which composite the item belongs to
#   - rationale:  why this item is being reversed
#
# After reversal, a new column `<item>_R` is created. The original column
# is preserved unchanged.

reverse_spec <- tibble::tribble(
  ~item,             ~scale_max, ~construct,            ~rationale,
  
  # --- Punitiveness items ---
  "punishmore_1",    7,          "Punitiveness",         "'Current punishment is too high' — agreeing means LESS punitive",
  "parsimony_1",     7,          "Punitiveness",         "'Punishments should be as light as possible' — agreeing = LESS punitive",
  "parsimony_2",     7,          "Punitiveness",         "'State should not punish more than necessary' — agreeing = LESS punitive",
  
  # --- Emotion items ---
  "hate_2",          7,          "Hatred",               "'I am sympathetic toward criminals' — sympathy is opposite of hatred",
  
  # --- Hostile aggression / harshness items ---
  "degrad_1",        7,          "Degradation",          "'Prisoners deserve dignity and respect' — agreeing = LOW degradation endorsement",
  "prisonvi_1",      7,          "Prison Violence",      "'I feel bad when a prisoner is assaulted' — feeling bad = NOT tolerating violence",
  "prisonvi_2",      7,          "Prison Violence",      "'Sexual assault of prisoners is a serious problem' — viewing as serious = NOT tolerating",
  "harsh_2",         7,          "Harsh Conditions",     "'Prisoners deserve decent living conditions' — agreeing = LOW harshness endorsement",
  
  # --- Crime concerns items ---
  "fear_1",          7,          "Fear of Crime",        "'How safe do you feel walking at night' — feeling safe = LOW fear",
  
  # --- Personality: RWA ---
  "rwa_3",           7,          "RWA",                  "'Nothing wrong with premarital sex' — disagreeing = HIGH authoritarianism",
  
  # --- Personality: SDO (egalitarian items reversed) ---
  "sdo_3",           7,          "SDO",                  "'No one group should dominate' — agreeing = LOW SDO (egalitarian)",
  "sdo_4",           7,          "SDO",                  "'Bottom groups are just as deserving' — agreeing = LOW SDO (egalitarian)",
  "sdo_7",           7,          "SDO",                  "'Should equalize conditions for groups' — agreeing = LOW SDO (egalitarian)",
  "sdo_8",           7,          "SDO",                  "'All groups deserve equal chance' — agreeing = LOW SDO (egalitarian)",
  
  # --- Personality: Vengefulness ---
  "venge_1",         7,          "Vengefulness",         "'Not worth my time to pay back someone' — agreeing = LOW vengefulness",
  "venge_5",         7,          "Vengefulness",         "'I am not a vengeful person' — agreeing = LOW vengefulness",
  
  # --- Personality: Racial Resentment ---
  "Raceresent_1",    7,          "Racial Resentment",    "'Slavery created difficult conditions' — agreeing = LOW racial resentment",
  "Raceresent_4",    7,          "Racial Resentment",    "'Blacks have gotten less than they deserve' — agreeing = LOW racial resentment",
  
  # --- Process violations (exploratory) ---
  "dueprocess_2",    7,          "Due Process",          "'Every defendant deserves due process' — agreeing = NOT tolerating violations"
)

# Print the reverse-coding plan for transparency
cat("=== REVERSE-CODING PLAN ===\n")
cat(sprintf("%d items will be reverse-coded.\n", nrow(reverse_spec)))
cat("Each item X becomes X_R via formula: X_R = (scale_max + 1) - X\n\n")
print(as.data.frame(reverse_spec %>% select(item, construct, rationale)),
      row.names = FALSE)
cat("\n")
```

```{r phase2_apply}
# === APPLY REVERSE CODING ===
# Single pass through the metadata table — each item becomes a new column.
df_clean <- df

for (i in seq_len(nrow(reverse_spec))) {
  item     <- reverse_spec$item[i]
  scale_max <- reverse_spec$scale_max[i]
  
  # Standard reverse coding: new = (max + 1) - old
  new_col <- paste0(item, "_R")
  df_clean[[new_col]] <- (scale_max + 1) - df_clean[[item]]
}

cat(sprintf("Reverse coding complete. %d new '_R' columns created.\n\n",
            nrow(reverse_spec)))

# Sanity check: confirm reversal is correct for one item
cat("Sanity check (punishmore_1):\n")
sanity <- df_clean %>%
  select(punishmore_1, punishmore_1_R) %>%
  count(punishmore_1, punishmore_1_R) %>%
  mutate(check = paste0("8 - ", punishmore_1, " = ", punishmore_1_R))
print(as.data.frame(sanity), row.names = FALSE)
cat("\n")
```


# PHASE 3: Z-SCORE SENTENCING DECISIONS WITHIN VIGNETTE

Each participant read **one** of three randomly assigned vignettes. The vignettes
differ in features (relationship to victim, premeditation, circumstances) which
likely shift the *absolute* sentence length participants find appropriate.

To use sentencing decisions as a punitiveness measure that's comparable across
participants regardless of vignette, we **z-score within vignette**: each
participant's sentence is expressed as a deviation from their vignette's mean,
in units of that vignette's SD. This preserves *individual differences* while
removing vignette-specific shifts.

```{r phase3_vignette_check}
# Confirm vignette distribution and labels
cat("=== VIGNETTE DISTRIBUTION ===\n")
print(table(df_clean$vignette, useNA = "ifany"))

cat("\nVignette labels:\n")
cat("  1 = Stranger Felony-Murder (purse snatching gone wrong)\n")
cat("  2 = Domestic Violence Murder (restraining order violation)\n")
cat("  3 = Organized Crime Murder (car theft operation)\n\n")
```

```{r phase3_descriptives_raw}
# Report raw sentence descriptives per vignette BEFORE z-scoring,
# so the reader can see the absolute differences being adjusted for.
cat("=== SENTENCE DESCRIPTIVES BY VIGNETTE (raw) ===\n\n")

sentence_by_vignette <- df_clean %>%
  group_by(vignette) %>%
  summarise(
    n      = n(),
    mean   = mean(Sentence_1, na.rm = TRUE),
    sd     = sd(Sentence_1, na.rm = TRUE),
    median = median(Sentence_1, na.rm = TRUE),
    min    = min(Sentence_1, na.rm = TRUE),
    max    = max(Sentence_1, na.rm = TRUE),
    .groups = "drop"
  )
print(as.data.frame(sentence_by_vignette))
cat("\n")
```

```{r phase3_zscore}
# Z-score sentencing decision within each vignette condition
df_clean <- df_clean %>%
  group_by(vignette) %>%
  mutate(Sentence_z = as.numeric(scale(Sentence_1))) %>%
  ungroup()

# Verify: means should be ~0 and SDs should be ~1 within each vignette
cat("=== VERIFICATION: Sentence_z descriptives by vignette ===\n\n")
verification <- df_clean %>%
  group_by(vignette) %>%
  summarise(
    n        = n(),
    mean_z   = mean(Sentence_z, na.rm = TRUE),
    sd_z     = sd(Sentence_z, na.rm = TRUE),
    .groups  = "drop"
  )
print(as.data.frame(verification))
cat("\n(Means should be ~0, SDs should be ~1)\n\n")
```

# PHASE 4: COMPOSITE CONSTRUCTION WITH RELIABILITY

Composites are constructed by averaging items (after reverse coding). Cronbach's
alpha is reported for every multi-item composite. Single-item constructs (LWOP,
deathpenalty, Sentence) are left as-is.

The composites are organized into five clusters:

  1. **Punitiveness** (DV) — penological attitudes + policy + sentencing
  2. **Crime Concerns** — perceived crime rates + fear of crime
  3. **Emotions** — hatred + anger
  4. **Hostile Aggression** — exclusion, degradation, suffering, prison violence,
     harsh conditions, revenge
  5. **Personality / Ideology** — RWA, SDO, vengefulness, violence proneness,
     racial resentment, blood sports

A sixth set (Process Violations: due process + uncertain evidence) is computed
as **exploratory** — collected and reported but not part of the confirmatory
hypotheses.

```{r phase4_alpha_table_init}
# Initialize the master alpha table that we'll populate as we build composites
alpha_table <- data.frame(
  Construct = character(),
  Alpha     = numeric(),
  N_Items   = integer(),
  stringsAsFactors = FALSE
)

# Helper to register a composite into the alpha table
register_alpha <- function(name, result) {
  data.frame(
    Construct = name,
    Alpha     = round(result$alpha, 3),
    N_Items   = result$n_items,
    stringsAsFactors = FALSE
  )
}
```

## 4.1 Punitiveness (dependent variable)

```{r phase4_punitiveness}
cat("=== PUNITIVENESS COMPOSITES ===\n\n")

# --- Punish More (2 items, one reversed) ---
punishmore_items <- c("punishmore_1_R", "punishmore_2")
res <- compute_composite(df_clean, punishmore_items, "Punish More")
df_clean$punishmore_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Punish More", res))

# --- Parsimony (2 items, both reversed → higher = LESS parsimonious = MORE punitive) ---
parsimony_items <- c("parsimony_1_R", "parsimony_2_R")
res <- compute_composite(df_clean, parsimony_items, "Parsimony (reversed)")
df_clean$parsimony_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Parsimony", res))

# --- Three Strikes (2 items) ---
threestrikes_items <- c("threestrikes_1", "threestrikes_2")
res <- compute_composite(df_clean, threestrikes_items, "Three Strikes")
df_clean$threestrikes_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Three Strikes", res))

# --- Single-item measures (no alpha to compute) ---
cat(sprintf("  %-32s Single item\n", "LWOP"))
cat(sprintf("  %-32s Single item\n", "Death Penalty"))
cat(sprintf("  %-32s Z-scored within vignette\n", "Sentence"))

# --- 8-item attitudes/policy composite (excludes sentence) ---
# This is the "battery-only" punitiveness measure
punitiveness_items <- c("punishmore_1_R", "punishmore_2",
                        "parsimony_1_R", "parsimony_2_R",
                        "threestrikes_1", "threestrikes_2",
                        "LWOP", "deathpenalty")
res <- compute_composite(df_clean, punitiveness_items, "Punitiveness (8 items)")
df_clean$punitiveness_8item   <- res$composite
df_clean$punitiveness_8item_z <- as.numeric(scale(df_clean$punitiveness_8item))
alpha_table <- rbind(alpha_table, register_alpha("Punitiveness (8 items)", res))

# --- Punitiveness AGGREGATE (the primary DV) ---
# Combines the standardized 8-item attitudes composite with the z-scored
# sentencing decision. This is what enters all the H1/H2/H3 tests.
df_clean$punitiveness_agg <- rowMeans(
  cbind(df_clean$punitiveness_8item_z, df_clean$Sentence_z),
  na.rm = TRUE
)

cat(sprintf("  %-32s Z-scored 8-item + Sentence_z (mean)\n\n", "Punitiveness Aggregate (DV)"))

# Per Dan's review (Comment 312): also compute alpha including the sentencing
# decision. We z-score sentence to put it on the same scale, then re-compute
# alpha across the 9-item composite.
puni9_data <- df_clean %>%
  select(all_of(punitiveness_items), Sentence_z) %>%
  mutate(across(all_of(punitiveness_items), ~ as.numeric(scale(.))))
alpha_with_sentence <- psych::alpha(puni9_data, check.keys = FALSE)$total$raw_alpha

cat(sprintf("Alpha with sentencing included (9 items, all standardized): α = %.3f\n\n",
            alpha_with_sentence))
alpha_table <- rbind(alpha_table,
                     data.frame(Construct = "Punitiveness (9 items, w/ sentence)",
                                Alpha     = round(alpha_with_sentence, 3),
                                N_Items   = 9))
```

## 4.2 Crime Concerns cluster

```{r phase4_crime_concerns}
cat("=== CRIME CONCERNS CLUSTER ===\n\n")

# --- Perceived Crime Rates (2 items) ---
rates_items <- c("rates_1", "rates_2")
res <- compute_composite(df_clean, rates_items, "Perceived Crime Rates")
df_clean$crime_rates_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Perceived Crime Rates", res))

# --- Fear of Crime (3 items, one reversed) ---
fear_items <- c("fear_1_R", "fear_2", "fear_3")
res <- compute_composite(df_clean, fear_items, "Fear of Crime")
df_clean$fear_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Fear of Crime", res))

# --- Crime Concerns aggregate (cluster-level composite) ---
crime_concerns_items <- c("rates_1", "rates_2", "fear_1_R", "fear_2", "fear_3")
res <- compute_composite(df_clean, crime_concerns_items, "Crime Concerns Cluster")
df_clean$crime_concerns_agg <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Crime Concerns Cluster", res))
cat("\n")
```

## 4.3 Emotions cluster

```{r phase4_emotions}
cat("=== EMOTIONS CLUSTER ===\n\n")

# --- Hatred (3 items, one reversed) ---
hatred_items <- c("hate_1", "hate_2_R", "hate_3")
res <- compute_composite(df_clean, hatred_items, "Hatred")
df_clean$hatred_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Hatred", res))

# --- Anger (2 items) ---
anger_items <- c("anger_1", "anger_2")
res <- compute_composite(df_clean, anger_items, "Anger")
df_clean$anger_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Anger", res))

# --- Emotions aggregate ---
emotions_items <- c("hate_1", "hate_2_R", "hate_3", "anger_1", "anger_2")
res <- compute_composite(df_clean, emotions_items, "Emotions Cluster")
df_clean$emotions_agg <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Emotions Cluster", res))
cat("\n")
```

## 4.4 Hostile Aggression cluster

The central correlate of interest. Six constructs that capture hostile,
retaliatory, and degradation-oriented dispositions toward people convicted of
crimes.

```{r phase4_hostile_aggression}
cat("=== HOSTILE AGGRESSION CLUSTER ===\n\n")

# --- Social Exclusion (3 items) ---
exclusion_items <- c("exclusion_1", "exclusion_2", "exclusion_3")
res <- compute_composite(df_clean, exclusion_items, "Social Exclusion")
df_clean$exclusion_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Social Exclusion", res))

# --- Degradation (3 items, one reversed) ---
degradation_items <- c("degrad_1_R", "degrad_2", "degrad_3")
res <- compute_composite(df_clean, degradation_items, "Degradation")
df_clean$degradation_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Degradation", res))

# --- Infliction of Suffering (2 items) ---
suffering_items <- c("suffer_1", "suffer_2")
res <- compute_composite(df_clean, suffering_items, "Infliction of Suffering")
df_clean$suffering_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Infliction of Suffering", res))

# --- Prison Violence Tolerance (2 items, both reversed) ---
prisonvi_items <- c("prisonvi_1_R", "prisonvi_2_R")
res <- compute_composite(df_clean, prisonvi_items, "Prison Violence Tolerance")
df_clean$prisonvi_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Prison Violence Tolerance", res))

# --- Harsh Prison Conditions (3 items, one reversed) ---
harsh_items <- c("harsh_1", "harsh_2_R", "harsh_3")
res <- compute_composite(df_clean, harsh_items, "Harsh Prison Conditions")
df_clean$harsh_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Harsh Prison Conditions", res))

# --- Revenge (3 items) ---
revenge_items <- c("revenge_1", "revenge_2", "revenge_3")
res <- compute_composite(df_clean, revenge_items, "Revenge")
df_clean$revenge_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Revenge", res))

# --- Hostile Aggression aggregate (cluster-level) ---
hostile_agg_items <- c(exclusion_items, degradation_items, suffering_items,
                       prisonvi_items, harsh_items, revenge_items)
res <- compute_composite(df_clean, hostile_agg_items, "Hostile Aggression Cluster")
df_clean$hostile_agg <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Hostile Aggression Cluster", res))
cat("\n")
```

## 4.5 Personality / Ideology cluster

```{r phase4_personality}
cat("=== PERSONALITY / IDEOLOGY CLUSTER ===\n\n")

# --- Right-Wing Authoritarianism (5 items, one reversed) ---
rwa_items <- c("rwa_1", "rwa_2", "rwa_3_R", "rwa_4", "rwa_5")
res <- compute_composite(df_clean, rwa_items, "RWA")
df_clean$rwa_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("RWA", res))

# --- Social Dominance Orientation (8 items, four reversed) ---
sdo_items <- c("sdo_1", "sdo_2", "sdo_3_R", "sdo_4_R",
               "sdo_5", "sdo_6", "sdo_7_R", "sdo_8_R")
res <- compute_composite(df_clean, sdo_items, "SDO")
df_clean$sdo_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("SDO", res))

# --- Vengefulness (5 items, two reversed) ---
venge_items <- c("venge_1_R", "venge_2", "venge_3", "venge_4", "venge_5_R")
res <- compute_composite(df_clean, venge_items, "Vengefulness")
df_clean$venge_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Vengefulness", res))

# --- Violence Proneness (4 items) ---
vprone_items <- c("vprone_1", "vprone_2", "vprone_3", "vprone_4")
res <- compute_composite(df_clean, vprone_items, "Violence Proneness")
df_clean$vprone_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Violence Proneness", res))

# --- Racial Resentment (4 items, two reversed) ---
raceresent_items <- c("Raceresent_1_R", "Raceresent_2", "Raceresent_3", "Raceresent_4_R")
res <- compute_composite(df_clean, raceresent_items, "Racial Resentment")
df_clean$raceresent_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Racial Resentment", res))

# --- Blood Sports (4 items: Boxing, MMA/UFC, Hunting, Wrestling) ---
# TV_4 = Boxing, TV_6 = MMA/UFC, TV_8 = Hunting, TV_9 = Wrestling
# Six other TV items (comedy, swimming, basketball, soccer, news, ice skating)
# are NOT included — they're collected for comparison only.
bloodsports_items <- c("TV_4", "TV_6", "TV_8", "TV_9")
res <- compute_composite(df_clean, bloodsports_items, "Blood Sports")
df_clean$bloodsports_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Blood Sports", res))

# --- Personality aggregate (all personality measures combined) ---
personality_items <- c(rwa_items, sdo_items, venge_items, vprone_items,
                       raceresent_items, bloodsports_items)
res <- compute_composite(df_clean, personality_items, "Personality Cluster")
df_clean$personality_agg <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Personality Cluster", res))
cat("\n")
```

## 4.6 Process Violations cluster (exploratory)

The Process Violations cluster (due process + uncertain evidence) was collected
but **dropped from confirmatory analyses per pre-registration**. We compute the
composites here so they're available for exploratory analyses in Phase 9.

```{r phase4_process}
cat("=== PROCESS VIOLATIONS (EXPLORATORY ONLY) ===\n\n")

# --- Due Process (3 items, one reversed) ---
dueprocess_items <- c("dueprocess_1", "dueprocess_2_R", "dueprocess_3")
res <- compute_composite(df_clean, dueprocess_items, "Due Process Violations")
df_clean$dueprocess_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Due Process Violations*", res))

# --- Uncertain Evidence (4 items) ---
uncertain_items <- c("uncertain_1", "uncertain_2", "uncertain_3", "uncertain_4")
res <- compute_composite(df_clean, uncertain_items, "Uncertain Evidence")
df_clean$uncertain_comp <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Uncertain Evidence*", res))

# --- Process Violations aggregate ---
process_items <- c(dueprocess_items, uncertain_items)
res <- compute_composite(df_clean, process_items, "Process Violations Cluster")
df_clean$process_agg <- res$composite
alpha_table <- rbind(alpha_table, register_alpha("Process Violations Cluster*", res))

cat("\n* Exploratory only (cluster dropped from confirmatory analyses)\n\n")
```

## 4.7 Complete reliability table

```{r phase4_alpha_summary}
cat("=== COMPLETE RELIABILITY TABLE ===\n\n")
print(alpha_table, row.names = FALSE)

# Flag low reliability composites for transparent reporting
low_alpha <- alpha_table %>% filter(Alpha < 0.60)
if (nrow(low_alpha) > 0) {
  cat("\nWARNING: Composites with α < .60 (interpret with caution):\n")
  print(as.data.frame(low_alpha), row.names = FALSE)
}
cat("\n")
```


# PHASE 5: DEMOGRAPHICS AND DESCRIPTIVE STATISTICS

## 5.1 Sample demographics

```{r phase5_demographics}
cat("=== SAMPLE DEMOGRAPHICS ===\n\n")

# --- Age ---
cat("AGE:\n")
cat(sprintf("  Mean: %.1f years\n", mean(df_clean$age_1, na.rm = TRUE)))
cat(sprintf("  SD:   %.1f\n", sd(df_clean$age_1, na.rm = TRUE)))
cat(sprintf("  Range: %d–%d\n\n",
            min(df_clean$age_1, na.rm = TRUE),
            max(df_clean$age_1, na.rm = TRUE)))

# --- Gender (1=Male, 2=Female, 3=Other) ---
cat("GENDER:\n")
gender_table <- df_clean %>%
  count(gender) %>%
  mutate(
    Gender_Label = case_when(
      gender == 1 ~ "Male",
      gender == 2 ~ "Female",
      gender == 3 ~ "Other",
      TRUE        ~ "Missing"),
    Percent = sprintf("%.1f%%", n / sum(n) * 100)
  ) %>%
  select(Gender_Label, n, Percent)
print(as.data.frame(gender_table), row.names = FALSE)
cat("\n")

# --- Race / Ethnicity ---
cat("RACE / ETHNICITY:\n")
race_table <- df_clean %>%
  count(race) %>%
  mutate(
    Race_Label = case_when(
      race == 1 ~ "White, non-Hispanic",
      race == 2 ~ "Black / African-American",
      race == 3 ~ "Asian",
      race == 4 ~ "Hispanic / Latino",
      race == 5 ~ "Middle Eastern",
      race == 6 ~ "Native American",
      race == 7 ~ "Pacific Islander",
      TRUE      ~ "Missing / Other"),
    Percent = sprintf("%.1f%%", n / sum(n) * 100)
  ) %>%
  select(Race_Label, n, Percent) %>%
  arrange(desc(n))
print(as.data.frame(race_table), row.names = FALSE)
cat("\n")

# --- Education ---
cat("EDUCATION:\n")
education_table <- df_clean %>%
  count(education) %>%
  mutate(
    Education_Label = case_when(
      education == 1 ~ "Did not graduate high school",
      education == 2 ~ "High school only",
      education == 3 ~ "1–2 years post-HS",
      education == 4 ~ "3–4 years post-HS",
      education == 5 ~ "5–6 years post-HS",
      education == 6 ~ "More than 6 years post-HS",
      TRUE           ~ "Missing"),
    Percent = sprintf("%.1f%%", n / sum(n) * 100)
  ) %>%
  select(Education_Label, n, Percent)
print(as.data.frame(education_table), row.names = FALSE)
cat("\n")

# --- Political Orientation (1=Strongly liberal, 7=Strongly conservative) ---
cat("POLITICAL ORIENTATION:\n")
cat(sprintf("  Mean:   %.2f (1=Strongly liberal, 7=Strongly conservative)\n",
            mean(df_clean$politid, na.rm = TRUE)))
cat(sprintf("  SD:     %.2f\n", sd(df_clean$politid, na.rm = TRUE)))
cat(sprintf("  Median: %.0f\n\n", median(df_clean$politid, na.rm = TRUE)))
```

## 5.2 Construct descriptive statistics

```{r phase5_descriptives}
cat("=== CONSTRUCT DESCRIPTIVE STATISTICS ===\n")
cat("(Pct columns show proportions below/at/above the scale midpoint)\n\n")

# Variables to describe — every composite + key single items + aggregates
construct_vars_desc <- c(
  # Punitiveness
  "punishmore_comp", "parsimony_comp", "threestrikes_comp",
  "LWOP", "deathpenalty", "Sentence_1", "punitiveness_agg",
  # Crime Concerns
  "crime_rates_comp", "fear_comp", "crime_concerns_agg",
  # Emotions
  "hatred_comp", "anger_comp", "emotions_agg",
  # Hostile Aggression
  "exclusion_comp", "degradation_comp", "suffering_comp",
  "prisonvi_comp", "harsh_comp", "revenge_comp", "hostile_agg",
  # Personality
  "rwa_comp", "sdo_comp", "venge_comp", "vprone_comp",
  "raceresent_comp", "bloodsports_comp", "personality_agg",
  # Process (exploratory)
  "dueprocess_comp", "uncertain_comp", "process_agg"
)

# Function to compute the percentage below/at/above scale midpoint
compute_midpoint_pcts <- function(x, midpoint) {
  n_valid <- sum(!is.na(x))
  c(Pct_Below = round(sum(x < midpoint, na.rm = TRUE) / n_valid, 2),
    Pct_At    = round(sum(x == midpoint, na.rm = TRUE) / n_valid, 2),
    Pct_Above = round(sum(x > midpoint, na.rm = TRUE) / n_valid, 2))
}

# Build descriptives row by row
descriptives_list <- lapply(construct_vars_desc, function(var) {
  x <- df_clean[[var]]

  # Determine appropriate midpoint per variable
  midpoint <- if (var == "Sentence_1") {
    25     # 0–50 scale
  } else if (var == "punitiveness_agg") {
    0      # standardized DV
  } else {
    4      # standard 7-point Likert midpoint
  }

  data.frame(
    Variable  = var,
    Mean      = round(mean(x, na.rm = TRUE), 2),
    SD        = round(sd(x, na.rm = TRUE), 2),
    Median    = round(median(x, na.rm = TRUE), 2),
    Min       = round(min(x, na.rm = TRUE), 2),
    Max       = round(max(x, na.rm = TRUE), 2),
    N         = sum(!is.na(x)),
    Pct_Below = compute_midpoint_pcts(x, midpoint)["Pct_Below"],
    Pct_At    = compute_midpoint_pcts(x, midpoint)["Pct_At"],
    Pct_Above = compute_midpoint_pcts(x, midpoint)["Pct_Above"],
    row.names = NULL
  )
})

descriptives_table <- do.call(rbind, descriptives_list)
print(descriptives_table, row.names = FALSE)
cat("\n")
```

## 5.3 Sentence distribution

Per Dan's Comment 435: report sentencing means by vignette and compare to the
25-year anchor mentioned in the instructions ("most sentences range from 20-30
years").

```{r phase5_sentence}
cat("=== SENTENCING DECISION DISTRIBUTION ===\n\n")

# Overall summary
cat("Overall:\n")
cat(sprintf("  Mean:   %.1f years\n",  mean(df_clean$Sentence_1, na.rm = TRUE)))
cat(sprintf("  SD:     %.1f\n",        sd(df_clean$Sentence_1,   na.rm = TRUE)))
cat(sprintf("  Median: %.0f\n",        median(df_clean$Sentence_1, na.rm = TRUE)))
cat(sprintf("  %% at maximum (50 yrs): %.1f%%\n",
            mean(df_clean$Sentence_1 == 50, na.rm = TRUE) * 100))
cat(sprintf("  %% at 0 years:          %.1f%%\n\n",
            mean(df_clean$Sentence_1 == 0, na.rm = TRUE) * 100))

# Per-vignette summary with comparison to 25-year anchor
cat("By Vignette (with comparison to 25-year anchor):\n")
sentence_vs_anchor <- df_clean %>%
  group_by(vignette) %>%
  summarise(
    n            = n(),
    mean         = round(mean(Sentence_1, na.rm = TRUE), 2),
    sd           = round(sd(Sentence_1,   na.rm = TRUE), 2),
    median       = median(Sentence_1, na.rm = TRUE),
    diff_25yr    = round(mean(Sentence_1, na.rm = TRUE) - 25, 2),
    pct_above_25 = round(mean(Sentence_1 > 25, na.rm = TRUE) * 100, 1),
    .groups      = "drop"
  )
print(as.data.frame(sentence_vs_anchor), row.names = FALSE)

# One-sample t-tests against the 25-year anchor for each vignette
cat("\nOne-sample t-tests vs. 25-year anchor:\n")
for (v in 1:3) {
  vd <- df_clean %>% filter(vignette == v) %>% pull(Sentence_1)
  tt <- t.test(vd, mu = 25)
  cat(sprintf("  Vignette %d: M = %.2f, t(%d) = %.2f, p %s\n",
              v, mean(vd, na.rm = TRUE), tt$parameter,
              tt$statistic, format_p(tt$p.value)))
}
cat("\n")
```

## 5.4 Sentence histograms

```{r phase5_histograms}
# --- Overall histogram ---
p_hist_overall <- ggplot(df_clean, aes(x = Sentence_1)) +
  geom_histogram(binwidth = 5, boundary = 0,
                 fill = "#3B6FA1", color = "white", linewidth = 0.4, alpha = 0.92) +
  geom_vline(xintercept = 25, color = "#C0392B",
             linetype = "dashed", linewidth = 0.7) +
  annotate("text", x = 25, y = Inf, label = "25-year anchor",
           hjust = -0.1, vjust = 2, color = "#C0392B",
           size = 3.8, fontface = "italic") +
  scale_x_continuous(breaks = seq(0, 50, 5),
                     expand = expansion(mult = c(0.01, 0.02))) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08))) +
  labs(title = "Distribution of Sentencing Decisions",
       x = "Years in prison", y = "Count") +
  theme_minimal(base_size = 13) +
  theme(
    plot.title         = element_text(hjust = 0.5, face = "bold",
                                      size = 15, margin = margin(b = 14)),
    axis.title.x       = element_text(margin = margin(t = 10)),
    axis.title.y       = element_text(margin = margin(r = 10)),
    axis.text          = element_text(color = "grey25"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    panel.grid.major.y = element_line(color = "grey90", linewidth = 0.4),
    plot.margin        = margin(t = 16, r = 18, b = 12, l = 12)
  )

ggsave(file.path(OUT_FIGURES, "sentence_histogram_overall.png"),
       p_hist_overall, width = 7.5, height = 4.5, dpi = 300)
cat("Saved: sentence_histogram_overall.png\n")

# --- By-vignette histogram (faceted) ---
vignette_labels <- c("1" = "Stranger Felony-Murder",
                     "2" = "Domestic Violence Murder",
                     "3" = "Organized Crime Murder")

p_hist_facet <- ggplot(df_clean, aes(x = Sentence_1, fill = factor(vignette))) +
  geom_histogram(binwidth = 5, color = "white", alpha = 0.8) +
  geom_vline(xintercept = 25, color = "red", linetype = "dashed", linewidth = 0.7) +
  scale_x_continuous(breaks = seq(0, 50, 10)) +
  scale_fill_brewer(palette = "Set2", guide = "none") +
  facet_wrap(~ vignette, ncol = 1,
             labeller = as_labeller(vignette_labels)) +
  labs(title = "Sentencing Decisions by Vignette",
       subtitle = "Red dashed line = 25-year anchor",
       x = "Years in prison", y = "Count") +
  theme_minimal(base_size = 12)

ggsave(file.path(OUT_FIGURES, "sentence_histogram_by_vignette.png"),
       p_hist_facet, width = 8, height = 9, dpi = 150)
cat("Saved: sentence_histogram_by_vignette.png\n\n")

# Display in knit
print(p_hist_overall)
print(p_hist_facet)
```

# PHASE 6: CONFIRMATORY HYPOTHESIS TESTS

This phase tests the three pre-registered hypotheses:

  - **H1**: Punitiveness positively correlated with all correlate measures
  - **H2**: Hostile aggression > Crime concerns in predicting punitiveness
  - **H3**: Most correlate measures positively intercorrelated

## 6.1 H1: Punitiveness correlates with all measures

### 6.1a Cluster-level correlations

```{r phase6_h1_cluster}
cat("=== H1: CLUSTER-LEVEL CORRELATIONS ===\n\n")

cluster_vars <- c("punitiveness_agg", "crime_concerns_agg", "process_agg",
                  "emotions_agg", "hostile_agg", "personality_agg")

# Display labels for clean output
cluster_labels <- c(
  "punitiveness_agg"   = "Punitiveness",
  "crime_concerns_agg" = "Crime Concerns",
  "process_agg"        = "Process Violations",
  "emotions_agg"       = "Emotions",
  "hostile_agg"        = "Hostile Aggression",
  "personality_agg"    = "Personality"
)

# Compute correlation matrix
cluster_data <- df_clean %>% select(all_of(cluster_vars)) %>% na.omit()
cluster_cor  <- corr.test(cluster_data, use = "pairwise", method = "pearson")

# Apply clean labels
colnames(cluster_cor$r) <- rownames(cluster_cor$r) <- cluster_labels[colnames(cluster_cor$r)]
colnames(cluster_cor$p) <- rownames(cluster_cor$p) <- cluster_labels[colnames(cluster_cor$p)]

cat("Correlation matrix (r):\n")
print(round(cluster_cor$r, 2))

cat("\nP-values:\n")
print(round(cluster_cor$p, 4))

# Punitiveness vs. each cluster (highlighted output)
cat("\n--- Punitiveness vs. Each Cluster ---\n")
for (var in cluster_vars[-1]) {
  r_val <- cor(df_clean$punitiveness_agg, df_clean[[var]],
               use = "pairwise.complete.obs")
  test  <- cor.test(df_clean$punitiveness_agg, df_clean[[var]])
  cat(sprintf("  %-22s r = %.2f, p %s\n",
              cluster_labels[var], r_val, format_p(test$p.value)))
}
cat("\n")
```

### 6.1b Construct-level correlations

```{r phase6_h1_construct}
cat("=== H1: CONSTRUCT-LEVEL CORRELATIONS ===\n\n")

# All 18 correlate constructs + punitiveness DV
construct_vars_h1 <- c(
  "punitiveness_agg",
  # Crime Concerns
  "crime_rates_comp", "fear_comp",
  # Process Violations (exploratory)
  "dueprocess_comp", "uncertain_comp",
  # Emotions
  "hatred_comp", "anger_comp",
  # Hostile Aggression
  "exclusion_comp", "degradation_comp", "suffering_comp",
  "prisonvi_comp", "harsh_comp", "revenge_comp",
  # Personality
  "rwa_comp", "sdo_comp", "venge_comp", "vprone_comp",
  "raceresent_comp", "bloodsports_comp"
)

# Display labels
construct_labels <- c(
  "punitiveness_agg" = "Punitiveness",
  "crime_rates_comp" = "Crime Rates",
  "fear_comp"        = "Fear",
  "dueprocess_comp"  = "Due Process",
  "uncertain_comp"   = "Uncertain Evidence",
  "hatred_comp"      = "Hatred",
  "anger_comp"       = "Anger",
  "exclusion_comp"   = "Exclusion",
  "degradation_comp" = "Degradation",
  "suffering_comp"   = "Suffering",
  "prisonvi_comp"    = "Prison Violence",
  "harsh_comp"       = "Harsh Conditions",
  "revenge_comp"     = "Revenge",
  "rwa_comp"         = "RWA",
  "sdo_comp"         = "SDO",
  "venge_comp"       = "Vengefulness",
  "vprone_comp"      = "Violence Proneness",
  "raceresent_comp"  = "Racial Resentment",
  "bloodsports_comp" = "Blood Sports"
)

# Compute correlation matrix
construct_data_h1 <- df_clean %>% select(all_of(construct_vars_h1)) %>% na.omit()
construct_cor     <- corr.test(construct_data_h1, use = "pairwise", method = "pearson")

# Apply clean labels
colnames(construct_cor$r) <- rownames(construct_cor$r) <- construct_labels[colnames(construct_cor$r)]
colnames(construct_cor$p) <- rownames(construct_cor$p) <- construct_labels[colnames(construct_cor$p)]

# Extract the punitiveness row
punitiveness_cors <- data.frame(
  Construct = colnames(construct_cor$r)[-1],
  r         = construct_cor$r["Punitiveness", -1],
  p         = construct_cor$p["Punitiveness", -1]
) %>%
  mutate(
    r         = round(r, 2),
    sig       = sig_stars(p),
    Direction = ifelse(r > 0, "Positive", "Negative")
  ) %>%
  arrange(desc(abs(r)))

cat("Punitiveness correlations with constructs (sorted by |r|):\n")
print(as.data.frame(punitiveness_cors), row.names = FALSE)

# H1 summary statistics
h1_n_positive     <- sum(punitiveness_cors$r > 0)
h1_n_total        <- nrow(punitiveness_cors)
h1_n_sig_positive <- sum(punitiveness_cors$r > 0 & punitiveness_cors$p < .05)

cat(sprintf("\n*** H1 SUMMARY (Construct Level) ***\n"))
cat(sprintf("  Positive correlations:      %d/%d (%.1f%%)\n",
            h1_n_positive, h1_n_total, h1_n_positive / h1_n_total * 100))
cat(sprintf("  Significant positive (p<.05): %d/%d (%.1f%%)\n",
            h1_n_sig_positive, h1_n_total, h1_n_sig_positive / h1_n_total * 100))
cat(sprintf("  H1 SUPPORTED: %s\n\n",
            ifelse(h1_n_positive == h1_n_total, "YES (all positive)", "PARTIAL")))
```

### 6.1c Item-level correlations

```{r phase6_h1_items}
cat("=== H1: ITEM-LEVEL CORRELATIONS WITH PUNITIVENESS ===\n\n")

# All 56 individual correlate items (after reverse coding)
all_items <- c(
  # Crime Concerns
  "rates_1", "rates_2", "fear_1_R", "fear_2", "fear_3",
  # Emotions
  "hate_1", "hate_2_R", "hate_3", "anger_1", "anger_2",
  # Hostile Aggression
  "exclusion_1", "exclusion_2", "exclusion_3",
  "degrad_1_R", "degrad_2", "degrad_3",
  "suffer_1", "suffer_2",
  "prisonvi_1_R", "prisonvi_2_R",
  "harsh_1", "harsh_2_R", "harsh_3",
  "revenge_1", "revenge_2", "revenge_3",
  # Personality
  "rwa_1", "rwa_2", "rwa_3_R", "rwa_4", "rwa_5",
  "sdo_1", "sdo_2", "sdo_3_R", "sdo_4_R",
  "sdo_5", "sdo_6", "sdo_7_R", "sdo_8_R",
  "venge_1_R", "venge_2", "venge_3", "venge_4", "venge_5_R",
  "vprone_1", "vprone_2", "vprone_3", "vprone_4",
  "Raceresent_1_R", "Raceresent_2", "Raceresent_3", "Raceresent_4_R",
  "TV_4", "TV_6", "TV_8", "TV_9"
)

# Compute correlations and p-values
item_cors  <- sapply(all_items, function(item)
  cor(df_clean$punitiveness_agg, df_clean[[item]], use = "pairwise.complete.obs"))

item_pvals <- sapply(all_items, function(item)
  cor.test(df_clean$punitiveness_agg, df_clean[[item]])$p.value)

item_cor_df <- data.frame(
  Item = all_items,
  r    = round(item_cors, 2),
  p    = item_pvals,
  sig  = sig_stars(item_pvals)
) %>% arrange(desc(abs(r)))

cat(sprintf("Item-level correlations with punitiveness: %d items\n",
            length(all_items)))
cat(sprintf("All positive: %s\n",
            ifelse(all(item_cors > 0), "YES", "NO")))
cat(sprintf("Proportion positive: %.1f%%\n",
            mean(item_cors > 0) * 100))
cat(sprintf("Significant (p < .05): %d/%d (%.1f%%)\n\n",
            sum(item_pvals < .05), length(item_pvals),
            sum(item_pvals < .05) / length(item_pvals) * 100))

print(item_cor_df, row.names = FALSE)
cat("\n")
```

### 6.1d Heatmap visualizations

```{r phase6_h1_heatmaps}
# --- Cluster-level heatmap ---
png(file.path(OUT_FIGURES, "heatmap_cluster_level.png"),
    width = 900, height = 900, res = 150)
corrplot(cluster_cor$r,
         method     = "color",
         type       = "upper",
         addCoef.col = "black",
         number.cex = 0.7,
         tl.col     = "black",
         tl.srt     = 45,
         tl.cex     = 0.75,
         p.mat      = cluster_cor$p,
         sig.level  = 0.05,
         insig      = "blank",
         title      = "Cluster-Level Correlations",
         mar        = c(0, 0, 2, 0))
dev.off()
cat("Saved: heatmap_cluster_level.png\n")

# --- Construct-level heatmap ---
png(file.path(OUT_FIGURES, "heatmap_construct_level.png"),
    width = 1400, height = 1400, res = 150)
corrplot(construct_cor$r,
         method     = "color",
         type       = "upper",
         addCoef.col = "black",
         number.cex = 0.45,
         tl.col     = "black",
         tl.srt     = 45,
         tl.cex     = 0.55,
         p.mat      = construct_cor$p,
         sig.level  = 0.05,
         insig      = "n",
         title      = "Construct-Level Correlations",
         mar        = c(0, 0, 2, 0))
dev.off()
cat("Saved: heatmap_construct_level.png\n\n")
```

### 6.1e 95% Confidence intervals for key cluster correlations

```{r phase6_h1_ci}
cat("=== 95% CIs FOR KEY CLUSTER CORRELATIONS ===\n\n")

ci_table <- data.frame(
  Comparison = c("Punitiveness — Hostile Aggression",
                 "Punitiveness — Crime Concerns",
                 "Punitiveness — Emotions",
                 "Punitiveness — Personality",
                 "Punitiveness — Process Violations"),
  r          = NA, CI_Lower = NA, CI_Upper = NA, p = NA
)

ci_table[1, 2:5] <- get_cor_ci("punitiveness_agg", "hostile_agg",         df_clean)
ci_table[2, 2:5] <- get_cor_ci("punitiveness_agg", "crime_concerns_agg",  df_clean)
ci_table[3, 2:5] <- get_cor_ci("punitiveness_agg", "emotions_agg",        df_clean)
ci_table[4, 2:5] <- get_cor_ci("punitiveness_agg", "personality_agg",     df_clean)
ci_table[5, 2:5] <- get_cor_ci("punitiveness_agg", "process_agg",         df_clean)

ci_table <- ci_table %>% mutate(across(where(is.numeric), ~ round(., 3)))
print(as.data.frame(ci_table), row.names = FALSE)
cat("\n")
```

## 6.2 H2: Hostile Aggression > Crime Concerns

This is the central confirmatory test. Steiger's Z compares two **dependent**
correlations (because both correlations share the punitiveness variable).

```{r phase6_h2_steiger}
cat("=== H2: HOSTILE AGGRESSION > CRIME CONCERNS ===\n\n")

# The three correlations we need for Steiger's test
r_hostile       <- cor(df_clean$punitiveness_agg, df_clean$hostile_agg,
                       use = "pairwise.complete.obs")
r_crime         <- cor(df_clean$punitiveness_agg, df_clean$crime_concerns_agg,
                       use = "pairwise.complete.obs")
r_hostile_crime <- cor(df_clean$hostile_agg,      df_clean$crime_concerns_agg,
                       use = "pairwise.complete.obs")

cat(sprintf("r(Punitiveness, Hostile Aggression) = %.3f\n", r_hostile))
cat(sprintf("r(Punitiveness, Crime Concerns)     = %.3f\n", r_crime))
cat(sprintf("r(Hostile, Crime)                   = %.3f\n\n", r_hostile_crime))

# Steiger's Z test using cocor for dependent overlapping correlations
n_steiger <- sum(complete.cases(df_clean[, c("punitiveness_agg",
                                              "hostile_agg",
                                              "crime_concerns_agg")]))
steiger_result <- cocor.dep.groups.overlap(
  r.jk = r_hostile,         # r(punitiveness, hostile)
  r.jh = r_crime,           # r(punitiveness, crime concerns)
  r.kh = r_hostile_crime,   # r(hostile, crime concerns)
  n    = n_steiger,
  test = "steiger1980"
)

cat("Steiger's Z-test for dependent correlations:\n")
cat(sprintf("  N = %d\n", n_steiger))
cat(sprintf("  r difference = %.3f − %.3f = %.3f\n",
            r_hostile, r_crime, r_hostile - r_crime))
print(steiger_result)

# Extract p-value
h2_pvalue <- steiger_result@steiger1980$p.value

cat(sprintf("\n*** H2 SUMMARY ***\n"))
cat(sprintf("  Hostile Aggression r = %.3f\n", r_hostile))
cat(sprintf("  Crime Concerns r     = %.3f\n", r_crime))
cat(sprintf("  Difference           = %.3f\n", r_hostile - r_crime))
cat(sprintf("  Steiger's Z p-value  = %.4f\n", h2_pvalue))
cat(sprintf("  H2 SUPPORTED:        %s\n\n",
            ifelse(r_hostile > r_crime & h2_pvalue < .05,
                   "YES (significantly greater)", "NO")))
```

### 6.2b Extended H2 — All clusters vs. Crime Concerns

```{r phase6_h2_extended}
cat("=== EXTENDED H2: ALL CLUSTERS vs. CRIME CONCERNS ===\n\n")

clusters_to_compare <- c("hostile_agg", "emotions_agg", "personality_agg", "process_agg")
cluster_labels_h2   <- c("Hostile Aggression", "Emotions", "Personality", "Process Violations")

# Collect results for export
extended_h2 <- data.frame(
  Cluster        = character(),
  r_Cluster      = numeric(),
  r_Crime        = numeric(),
  Difference     = numeric(),
  Steiger_p      = numeric(),
  Greater_p_lt05 = character(),
  stringsAsFactors = FALSE
)

for (i in seq_along(clusters_to_compare)) {
  r_other   <- cor(df_clean$punitiveness_agg, df_clean[[clusters_to_compare[i]]],
                   use = "pairwise.complete.obs")
  r_between <- cor(df_clean[[clusters_to_compare[i]]], df_clean$crime_concerns_agg,
                   use = "pairwise.complete.obs")

  steiger <- cocor.dep.groups.overlap(
    r.jk = r_other, r.jh = r_crime, r.kh = r_between,
    n    = n_steiger, test = "steiger1980"
  )

  cat(sprintf("%s vs Crime Concerns:\n", cluster_labels_h2[i]))
  cat(sprintf("  r(%s) = %.3f vs r(Crime) = %.3f\n",
              cluster_labels_h2[i], r_other, r_crime))
  cat(sprintf("  Difference = %.3f, p = %.4f\n",
              r_other - r_crime, steiger@steiger1980$p.value))
  cat(sprintf("  %s > Crime Concerns: %s\n\n", cluster_labels_h2[i],
              ifelse(steiger@steiger1980$p.value < .05 & r_other > r_crime,
                     "YES", "NO")))

  extended_h2 <- rbind(extended_h2, data.frame(
    Cluster        = cluster_labels_h2[i],
    r_Cluster      = round(r_other, 3),
    r_Crime        = round(r_crime, 3),
    Difference     = round(r_other - r_crime, 3),
    Steiger_p      = round(steiger@steiger1980$p.value, 4),
    Greater_p_lt05 = ifelse(steiger@steiger1980$p.value < .05 & r_other > r_crime,
                            "YES", "NO")
  ))
}
```

## 6.3 H3: Intercorrelations among correlates

```{r phase6_h3}
cat("=== H3: CORRELATES POSITIVELY INTERCORRELATED ===\n\n")

# All 18 correlate constructs (excluding the punitiveness DV)
correlate_vars <- c(
  "crime_rates_comp", "fear_comp",
  "dueprocess_comp", "uncertain_comp",
  "hatred_comp", "anger_comp",
  "exclusion_comp", "degradation_comp", "suffering_comp",
  "prisonvi_comp", "harsh_comp", "revenge_comp",
  "rwa_comp", "sdo_comp", "venge_comp", "vprone_comp",
  "raceresent_comp", "bloodsports_comp"
)

correlate_data <- df_clean %>% select(all_of(correlate_vars)) %>% na.omit()
correlate_cor  <- corr.test(correlate_data, use = "pairwise", method = "pearson")

# Apply clean labels
colnames(correlate_cor$r) <- rownames(correlate_cor$r) <- construct_labels[colnames(correlate_cor$r)]
colnames(correlate_cor$p) <- rownames(correlate_cor$p) <- construct_labels[colnames(correlate_cor$p)]

# Count correlations: extract upper triangle (off-diagonal, no double-counting)
n_vars            <- length(correlate_vars)
h3_n_correlations <- (n_vars * (n_vars - 1)) / 2
upper_r           <- correlate_cor$r[upper.tri(correlate_cor$r)]
upper_p           <- correlate_cor$p[upper.tri(correlate_cor$p)]

h3_n_positive     <- sum(upper_r > 0)
h3_n_sig          <- sum(upper_p < .05)
h3_n_sig_positive <- sum(upper_r > 0 & upper_p < .05)

cat(sprintf("Total off-diagonal correlations: %d\n", h3_n_correlations))
cat(sprintf("Positive correlations:           %d (%.1f%%)\n",
            h3_n_positive, h3_n_positive / h3_n_correlations * 100))
cat(sprintf("Significant (p < .05):           %d (%.1f%%)\n",
            h3_n_sig, h3_n_sig / h3_n_correlations * 100))
cat(sprintf("Significant AND positive:        %d (%.1f%%)\n",
            h3_n_sig_positive,
            h3_n_sig_positive / h3_n_correlations * 100))

cat(sprintf("\n*** H3 SUMMARY ***\n"))
cat(sprintf("  H3 SUPPORTED: %s\n\n",
            ifelse(h3_n_positive / h3_n_correlations > 0.5,
                   "YES (majority positive)", "NO")))

# Full intercorrelation heatmap
png(file.path(OUT_FIGURES, "heatmap_intercorrelations.png"),
    width = 1600, height = 1600, res = 150)
corrplot(correlate_cor$r,
         method     = "color",
         type       = "upper",
         addCoef.col = "black",
         number.cex = 0.45,
         tl.col     = "black",
         tl.srt     = 45,
         tl.cex     = 0.55,
         p.mat      = correlate_cor$p,
         sig.level  = 0.05,
         insig      = "n",
         title      = "Intercorrelations Among Correlate Constructs",
         mar        = c(0, 0, 2, 0))
dev.off()
cat("Saved: heatmap_intercorrelations.png\n\n")
```

## 6.4 FDR correction

Given the 18 simultaneous tests for H1, we apply False Discovery Rate
correction (Benjamini-Hochberg) to control the expected proportion of false
positives.

```{r phase6_fdr}
cat("=== FDR CORRECTION FOR MULTIPLE COMPARISONS ===\n\n")

# Apply FDR correction to all H1 construct-level p-values
all_pvalues  <- punitiveness_cors$p
fdr_adjusted <- p.adjust(all_pvalues, method = "fdr")

fdr_table <- punitiveness_cors %>%
  mutate(p_fdr   = round(fdr_adjusted, 4),
         sig_fdr = sig_stars(fdr_adjusted)) %>%
  select(Construct, r, p, sig, p_fdr, sig_fdr)

cat("H1 Results with FDR correction:\n")
print(as.data.frame(fdr_table), row.names = FALSE)

h1_n_sig_fdr <- sum(fdr_adjusted < .05 & punitiveness_cors$r > 0)
cat(sprintf("\nSignificant after FDR correction: %d/%d\n\n",
            h1_n_sig_fdr, nrow(fdr_table)))
```


# PHASE 7: CORRELATION TABLES FOR EXPORT

```{r phase7_tables}
cat("=== EXPORTING CORRELATION TABLES ===\n\n")

# --- Cluster-level table (printed) ---
cat("CLUSTER-LEVEL CORRELATION TABLE:\n\n")
cluster_table <- format_cor_table(cluster_cor$r, cluster_cor$p)
print(noquote(cluster_table))
cat("\nNote: * p < .05, ** p < .01, *** p < .001\n\n")

# --- Construct-level table (CSV only — too large to print) ---
write.csv(round(construct_cor$r, 2),
          file.path(OUT_TABLES, "correlation_table_construct_level.csv"))
cat("Saved: correlation_table_construct_level.csv\n")

write.csv(round(construct_cor$p, 4),
          file.path(OUT_TABLES, "pvalues_construct_level.csv"))
cat("Saved: pvalues_construct_level.csv\n")

# --- Intercorrelation table (correlates only, no DV) ---
correlate_table <- format_cor_table(correlate_cor$r, correlate_cor$p)
write.csv(correlate_table,
          file.path(OUT_TABLES, "correlation_table_intercorrelations.csv"))
cat("Saved: correlation_table_intercorrelations.csv\n\n")
```

# PHASE 8: VIGNETTE STABILITY SENSITIVITY ANALYSES

Although vignettes were randomly assigned, we test whether the substantive
findings depend on which vignette participants saw. Results that hold across
all three vignettes are more credible than those that emerge in only one.

## 8.1 Vignette effects on sentencing decisions

```{r phase8_anova}
cat("=== VIGNETTE EFFECTS ON SENTENCE ===\n\n")

# One-way ANOVA testing whether vignette affects raw sentence length
anova_model <- aov(Sentence_1 ~ factor(vignette), data = df_clean)
cat("One-way ANOVA: Sentence ~ Vignette\n")
print(summary(anova_model))

# Effect size (eta-squared)
eta_sq <- effectsize::eta_squared(anova_model)
cat("\nEffect size (eta-squared):\n")
print(as.data.frame(eta_sq))

# Tukey HSD post-hoc comparisons
cat("\nTukey HSD post-hoc comparisons:\n")
print(TukeyHSD(anova_model))
cat("\n")
```

## 8.2 Correlation stability across vignettes

```{r phase8_stability}
cat("=== CORRELATION STABILITY ACROSS VIGNETTES ===\n\n")

# Compute punitiveness correlations separately within each vignette
compute_vignette_cors <- function(data, vignette_num) {
  subset_data <- data %>% filter(vignette == vignette_num)
  sapply(correlate_vars, function(var)
    cor(subset_data$punitiveness_agg, subset_data[[var]],
        use = "pairwise.complete.obs"))
}

vignette_cors <- data.frame(
  Construct  = construct_labels[correlate_vars],
  Vignette_1 = compute_vignette_cors(df_clean, 1),
  Vignette_2 = compute_vignette_cors(df_clean, 2),
  Vignette_3 = compute_vignette_cors(df_clean, 3)
)

# Add overall correlation for comparison
vignette_cors$Overall <- sapply(correlate_vars, function(var)
  cor(df_clean$punitiveness_agg, df_clean[[var]],
      use = "pairwise.complete.obs"))

# Round
vignette_cors <- vignette_cors %>%
  mutate(across(where(is.numeric), ~ round(., 2)))

cat("Punitiveness correlations by vignette:\n")
print(as.data.frame(vignette_cors), row.names = FALSE)

# Compute range across vignettes — large ranges indicate instability
vignette_cors$Range <- apply(
  vignette_cors[, c("Vignette_1", "Vignette_2", "Vignette_3")],
  1, function(x) max(x) - min(x)
)

cat("\n\nLargest ranges (least stable correlations):\n")
print(as.data.frame(vignette_cors %>%
                    arrange(desc(Range)) %>%
                    head(5) %>%
                    select(Construct, Vignette_1, Vignette_2, Vignette_3, Range)),
      row.names = FALSE)
cat("\n")
```

## 8.3 Mixed-effects model with vignette random intercept

```{r phase8_mixed}
cat("=== MIXED-EFFECTS MODEL (Vignette as Random Intercept) ===\n\n")

# Model the central H2 relationship with vignette as a random intercept,
# allowing the baseline level of punitiveness to vary by vignette.
mixed_model <- lmer(punitiveness_agg ~ hostile_agg + (1 | vignette),
                    data = df_clean)

cat("Model: punitiveness_agg ~ hostile_agg + (1 | vignette)\n\n")
print(summary(mixed_model))

# Compare to pure fixed-effects model
fixed_model <- lm(punitiveness_agg ~ hostile_agg, data = df_clean)
cat(sprintf("\nFixed-effects R²: %.3f\n", summary(fixed_model)$r.squared))

# Intraclass correlation (proportion of variance attributable to vignette)
icc_result <- tryCatch({
  performance::icc(mixed_model)
}, warning = function(w) NULL, error = function(e) NULL)

if (is.null(icc_result) || is.atomic(icc_result)) {
  cat("ICC: Near zero (vignette accounts for negligible variance — singular fit)\n\n")
} else if (is.list(icc_result) && !is.null(icc_result$ICC_adjusted)) {
  cat(sprintf("ICC (Vignette): %.4f\n", icc_result$ICC_adjusted))
  cat("(Low ICC suggests vignette accounts for minimal variance)\n\n")
} else {
  cat("ICC: Could not be computed (likely near zero)\n\n")
}
```


# PHASE 9: EXPLORATORY ANALYSES

## 9.1 Process Violations cluster (dropped from confirmatory)

```{r phase9_process}
cat("=== PROCESS VIOLATIONS (EXPLORATORY) ===\n\n")

# Punitiveness correlations
process_cor_due <- cor.test(df_clean$punitiveness_agg, df_clean$dueprocess_comp)
process_cor_unc <- cor.test(df_clean$punitiveness_agg, df_clean$uncertain_comp)
process_cor_agg <- cor.test(df_clean$punitiveness_agg, df_clean$process_agg)

cat("Punitiveness correlations with Process Violations:\n")
cat(sprintf("  Due Process Violations:       r = %.2f, p %s\n",
            process_cor_due$estimate, format_p(process_cor_due$p.value)))
cat(sprintf("  Uncertain Evidence:           r = %.2f, p %s\n",
            process_cor_unc$estimate, format_p(process_cor_unc$p.value)))
cat(sprintf("  Process Violations Aggregate: r = %.2f, p %s\n\n",
            process_cor_agg$estimate, format_p(process_cor_agg$p.value)))

# Process Violations correlations with other clusters
cat("Process Violations correlations with other clusters:\n")
process_cluster_vars <- c("crime_concerns_agg", "emotions_agg",
                          "hostile_agg", "personality_agg")
for (var in process_cluster_vars) {
  r <- cor(df_clean$process_agg, df_clean[[var]],
           use = "pairwise.complete.obs")
  cat(sprintf("  %-22s r = %.2f\n", cluster_labels[var], r))
}
cat("\n")
```

## 9.2 Blood sports analysis

```{r phase9_blood_sports}
cat("=== BLOOD SPORTS ANALYSIS ===\n\n")

# Individual sport correlations with punitiveness
tv_sports <- c("TV_4", "TV_6", "TV_8", "TV_9")
tv_labels <- c("Boxing", "MMA / UFC", "Hunting", "Wrestling")

cat("Individual blood sport correlations with punitiveness:\n")
for (i in seq_along(tv_sports)) {
  r    <- cor(df_clean$punitiveness_agg, df_clean[[tv_sports[i]]],
              use = "pairwise.complete.obs")
  test <- cor.test(df_clean$punitiveness_agg, df_clean[[tv_sports[i]]])
  cat(sprintf("  %-12s r = %.2f, p %s\n",
              tv_labels[i], r, format_p(test$p.value)))
}

cat(sprintf("\n  Blood Sports Composite: r = %.2f\n",
            cor(df_clean$punitiveness_agg, df_clean$bloodsports_comp,
                use = "pairwise.complete.obs")))
cat("\n")
```

## 9.3 Revenge (beliefs) vs. Vengefulness (personality trait)

These two constructs capture related but distinct phenomena: Revenge is a
*belief* about whether punishment should serve retaliation, whereas
Vengefulness is a *personality trait* — a stable disposition. We test
whether they are statistically distinguishable predictors of punitiveness.

```{r phase9_revenge_vengefulness}
cat("=== REVENGE (beliefs) vs. VENGEFULNESS (trait) ===\n\n")

# Their correlation with each other
rev_veng_test <- cor.test(df_clean$revenge_comp, df_clean$venge_comp)
cat(sprintf("Correlation between constructs: r = %.3f, p %s\n",
            rev_veng_test$estimate, format_p(rev_veng_test$p.value)))
cat(sprintf("95%% CI: [%.3f, %.3f]\n\n",
            rev_veng_test$conf.int[1], rev_veng_test$conf.int[2]))

# Each with punitiveness
rev_pun_test  <- cor.test(df_clean$punitiveness_agg, df_clean$revenge_comp)
veng_pun_test <- cor.test(df_clean$punitiveness_agg, df_clean$venge_comp)

cat("Correlations with Punitiveness:\n")
cat(sprintf("  Revenge (beliefs):    r = %.3f, p %s\n",
            rev_pun_test$estimate,  format_p(rev_pun_test$p.value)))
cat(sprintf("  Vengefulness (trait): r = %.3f, p %s\n\n",
            veng_pun_test$estimate, format_p(veng_pun_test$p.value)))

# Steiger's Z to test whether one is a stronger predictor than the other
r_rev_veng <- cor(df_clean$revenge_comp, df_clean$venge_comp,
                  use = "pairwise.complete.obs")
n_rv       <- sum(complete.cases(df_clean[, c("punitiveness_agg",
                                              "revenge_comp",
                                              "venge_comp")]))
steiger_rv <- cocor.dep.groups.overlap(
  r.jk = rev_pun_test$estimate,
  r.jh = veng_pun_test$estimate,
  r.kh = r_rev_veng,
  n    = n_rv,
  test = "steiger1980"
)

cat("Steiger's Z (Revenge vs Vengefulness predicting Punitiveness):\n")
cat(sprintf("  Difference = %.3f, p = %.4f\n",
            rev_pun_test$estimate - veng_pun_test$estimate,
            steiger_rv@steiger1980$p.value))
cat(sprintf("  Revenge significantly stronger: %s\n\n",
            ifelse(steiger_rv@steiger1980$p.value < .05 &
                   rev_pun_test$estimate > veng_pun_test$estimate,
                   "YES", "NO")))
```

## 9.4 Exploratory factor analysis

Pre-registered EFA on the 18 correlate constructs to examine the latent
factor structure.

```{r phase9_efa}
cat("=== EXPLORATORY FACTOR ANALYSIS ===\n\n")

efa_vars <- correlate_vars
efa_data <- df_clean %>% select(all_of(efa_vars)) %>% na.omit()

cat(sprintf("Sample size for EFA:    N = %d\n", nrow(efa_data)))
cat(sprintf("Variables:              k = %d\n", ncol(efa_data)))
cat(sprintf("Subjects per variable:  %.1f\n\n", nrow(efa_data) / ncol(efa_data)))

# KMO Measure of Sampling Adequacy
kmo_result <- KMO(efa_data)
cat(sprintf("KMO Measure: %.3f\n", kmo_result$MSA))
cat("  (>.90 marvelous, >.80 meritorious, >.70 middling)\n\n")

# Parallel analysis to determine number of factors
cat("Parallel analysis (Horn's method):\n")
parallel_result <- fa.parallel(efa_data, fa = "fa", n.iter = 100, show.legend = FALSE)
cat("\n")

# Run EFA with the suggested number of factors
n_factors <- parallel_result$nfact
cat(sprintf("Running EFA with %d factors...\n\n", n_factors))
efa_result <- fa(efa_data, nfactors = n_factors, rotate = "varimax", fm = "ml")

# Loadings (cutoff 0.3 for clarity)
cat("Factor loadings (sorted, |loading| > 0.3 shown):\n")
print(efa_result$loadings, cutoff = 0.3, sort = TRUE)

cat("\nVariance explained:\n")
print(efa_result$Vaccounted)
cat("\n")

# Save scree plot
png(file.path(OUT_FIGURES, "efa_scree_plot.png"),
    width = 800, height = 600, res = 150)
plot(efa_result$values, type = "b",
     main = "EFA Scree Plot",
     xlab = "Factor", ylab = "Eigenvalue")
abline(h = 1, lty = 2, col = "red")
dev.off()
cat("Saved: efa_scree_plot.png\n\n")
```

## 9.5 Confirmatory factor analysis (pre-registered)

The five-factor CFA model corresponds to the theoretical clustering used
throughout the analysis.

```{r phase9_cfa}
cat("=== CONFIRMATORY FACTOR ANALYSIS ===\n\n")

# Theoretical five-factor model
cfa_model <- '
  CrimeConcerns     =~ crime_rates_comp + fear_comp
  ProcessViolations =~ dueprocess_comp + uncertain_comp
  Emotions          =~ hatred_comp + anger_comp
  HostileAggression =~ exclusion_comp + degradation_comp + suffering_comp +
                       prisonvi_comp + harsh_comp + revenge_comp
  Personality       =~ rwa_comp + sdo_comp + venge_comp + vprone_comp +
                       raceresent_comp + bloodsports_comp
'

# Fit with ML, fall back to MLR (robust) if non-convergence
cfa_fit <- tryCatch({
  cfa(cfa_model, data = df_clean, std.lv = TRUE)
}, error = function(e) {
  cat("Standard CFA did not converge. Trying robust estimator (MLR)...\n")
  cfa(cfa_model, data = df_clean, std.lv = TRUE, estimator = "MLR")
})

# Fit indices
cat("CFA Model fit indices:\n")
fit_indices <- fitMeasures(cfa_fit, c("chisq", "df", "pvalue",
                                       "cfi", "tli",
                                       "rmsea", "rmsea.ci.lower", "rmsea.ci.upper",
                                       "srmr"))
print(round(fit_indices, 3))

cat("\nInterpretation guidelines:\n")
cat("  CFI/TLI > .90 acceptable, > .95 good\n")
cat("  RMSEA < .08 acceptable, < .05 good\n")
cat("  SRMR  < .08 acceptable\n\n")

# Standardized factor loadings
cat("Standardized factor loadings:\n")
std_solution <- standardizedSolution(cfa_fit) %>%
  filter(op == "=~") %>%
  select(Factor = lhs, Indicator = rhs, Loading = est.std, p = pvalue) %>%
  mutate(Loading = round(Loading, 3),
         p       = round(p, 4))
print(as.data.frame(std_solution), row.names = FALSE)

# Factor correlations
cat("\nFactor correlations:\n")
factor_cors <- standardizedSolution(cfa_fit) %>%
  filter(op == "~~", lhs != rhs) %>%
  select(Factor1 = lhs, Factor2 = rhs, r = est.std) %>%
  mutate(r = round(r, 3))
print(as.data.frame(factor_cors), row.names = FALSE)
cat("\n")
```

## 9.6 Demographic moderators

Per Dan's Comment 433, we test whether punitiveness and sentencing decisions
interact with key demographics (age, gender, education, race, political
orientation).

```{r phase9_demographics}
cat("=== DEMOGRAPHIC MODERATORS ===\n\n")

# Recode political orientation into liberal / moderate / conservative
df_clean <- df_clean %>%
  mutate(political_group = case_when(
    politid <= 3 ~ "Liberal",
    politid == 4 ~ "Moderate",
    politid >= 5 ~ "Conservative"
  ))

# --- Political orientation ---
cat("POLITICAL ORIENTATION:\n")
political_cors <- df_clean %>%
  group_by(political_group) %>%
  summarise(
    n             = n(),
    mean_punitive = round(mean(punitiveness_agg, na.rm = TRUE), 3),
    r_hostile     = round(cor(punitiveness_agg, hostile_agg,
                              use = "pairwise.complete.obs"), 3),
    r_crime       = round(cor(punitiveness_agg, crime_concerns_agg,
                              use = "pairwise.complete.obs"), 3),
    r_personality = round(cor(punitiveness_agg, personality_agg,
                              use = "pairwise.complete.obs"), 3),
    .groups       = "drop"
  )
print(as.data.frame(political_cors), row.names = FALSE)
cat("\n")

# --- Gender ---
cat("GENDER:\n")
gender_cors <- df_clean %>%
  filter(gender %in% c(1, 2)) %>%
  mutate(gender_label = ifelse(gender == 1, "Male", "Female")) %>%
  group_by(gender_label) %>%
  summarise(
    n             = n(),
    mean_punitive = round(mean(punitiveness_agg, na.rm = TRUE), 3),
    r_hostile     = round(cor(punitiveness_agg, hostile_agg,
                              use = "pairwise.complete.obs"), 3),
    .groups       = "drop"
  )
print(as.data.frame(gender_cors), row.names = FALSE)

# T-test for gender difference in punitiveness
gender_ttest <- t.test(punitiveness_agg ~ gender,
                       data = df_clean %>% filter(gender %in% c(1, 2)))
cat(sprintf("\nGender difference: t(%.0f) = %.2f, p = %.3f\n",
            gender_ttest$parameter, gender_ttest$statistic,
            gender_ttest$p.value))

gender_d <- effectsize::cohens_d(punitiveness_agg ~ gender,
                                 data = df_clean %>% filter(gender %in% c(1, 2)))
cat(sprintf("Cohen's d = %.2f\n\n", gender_d$Cohens_d))

# --- Age (continuous correlation) ---
cat("AGE (continuous):\n")
age_pun_cor <- cor.test(df_clean$age_1, df_clean$punitiveness_agg)
cat(sprintf("  r(age, punitiveness)  = %.3f, p %s\n",
            age_pun_cor$estimate, format_p(age_pun_cor$p.value)))
age_sent_cor <- cor.test(df_clean$age_1, df_clean$Sentence_1)
cat(sprintf("  r(age, sentence)      = %.3f, p %s\n\n",
            age_sent_cor$estimate, format_p(age_sent_cor$p.value)))

# --- Education (continuous correlation) ---
cat("EDUCATION (continuous):\n")
edu_pun_cor <- cor.test(df_clean$education, df_clean$punitiveness_agg)
cat(sprintf("  r(education, punitiveness) = %.3f, p %s\n",
            edu_pun_cor$estimate, format_p(edu_pun_cor$p.value)))
edu_sent_cor <- cor.test(df_clean$education, df_clean$Sentence_1)
cat(sprintf("  r(education, sentence)     = %.3f, p %s\n\n",
            edu_sent_cor$estimate, format_p(edu_sent_cor$p.value)))

# --- Race (one-way ANOVA) ---
cat("RACE (ANOVA):\n")
race_anova <- aov(punitiveness_agg ~ factor(race), data = df_clean)
cat("  Punitiveness ~ Race:\n")
print(summary(race_anova))
race_anova_sent <- aov(Sentence_1 ~ factor(race), data = df_clean)
cat("\n  Sentence ~ Race:\n")
print(summary(race_anova_sent))
cat("\n")

# --- Political × Hostile interaction (continuous politid) ---
cat("POLITICAL × HOSTILE INTERACTION:\n")
political_int <- lm(punitiveness_agg ~ hostile_agg * politid, data = df_clean)
cat("Model: punitiveness_agg ~ hostile_agg * politid\n\n")
print(summary(political_int))
cat("\n")
```


# PHASE 10: MULTIPLE REGRESSION

## 10.1 All clusters predicting punitiveness

```{r phase10_clusters}
cat("=== ALL CLUSTERS PREDICTING PUNITIVENESS ===\n\n")

# Multi-cluster regression
cluster_reg <- lm(punitiveness_agg ~ crime_concerns_agg + process_agg +
                                     emotions_agg + hostile_agg + personality_agg,
                  data = df_clean)

cat("Model: Punitiveness ~ Crime + Process + Emotions + Hostile + Personality\n\n")
print(summary(cluster_reg))

# Standardized betas (re-fit with scaled variables)
cat("\nStandardized coefficients (β):\n")
cluster_reg_std <- lm(scale(punitiveness_agg) ~
                        scale(crime_concerns_agg) + scale(process_agg) +
                        scale(emotions_agg) + scale(hostile_agg) +
                        scale(personality_agg),
                      data = df_clean)
betas <- round(coef(cluster_reg_std)[-1], 3)
print(betas)

# Unique variance via Type-III ANOVA
cat("\nUnique variance explained (Type-III SS / total SS):\n")
anova_type3 <- Anova(cluster_reg, type = 3)
ss_total    <- sum(anova_type3$`Sum Sq`)
unique_var  <- anova_type3$`Sum Sq` / ss_total
names(unique_var) <- rownames(anova_type3)
print(round(unique_var[-1], 4))  # exclude intercept
cat("\n")
```

## 10.2 All constructs predicting punitiveness

```{r phase10_constructs}
cat("=== ALL CONSTRUCTS PREDICTING PUNITIVENESS ===\n\n")

construct_reg <- lm(punitiveness_agg ~
                      crime_rates_comp + fear_comp +
                      dueprocess_comp + uncertain_comp +
                      hatred_comp + anger_comp +
                      exclusion_comp + degradation_comp + suffering_comp +
                      prisonvi_comp + harsh_comp + revenge_comp +
                      rwa_comp + sdo_comp + venge_comp + vprone_comp +
                      raceresent_comp + bloodsports_comp,
                    data = df_clean)

cat("Full construct-level model summary:\n")
print(summary(construct_reg))

# Significant predictors
cat("\nSignificant predictors (p < .05), sorted by |β|:\n")
construct_summary <- broom::tidy(construct_reg) %>%
  filter(p.value < .05, term != "(Intercept)") %>%
  arrange(desc(abs(estimate)))
print(as.data.frame(construct_summary), row.names = FALSE)
cat("\n")
```

## 10.3 Hierarchical regression (incremental R²)

This regression progressively adds clusters in a theoretically motivated order:
crime concerns first (the prosocially "legitimate" predictor), then process
violations, hostile aggression, emotions, and finally personality. We track
ΔR² at each step to quantify how much each cluster adds beyond the previous.

```{r phase10_hierarchical}
cat("=== HIERARCHICAL REGRESSION ===\n\n")

step1 <- lm(punitiveness_agg ~ crime_concerns_agg,                                    data = df_clean)
step2 <- lm(punitiveness_agg ~ crime_concerns_agg + process_agg,                      data = df_clean)
step3 <- lm(punitiveness_agg ~ crime_concerns_agg + process_agg + hostile_agg,        data = df_clean)
step4 <- lm(punitiveness_agg ~ crime_concerns_agg + process_agg + hostile_agg + emotions_agg, data = df_clean)
step5 <- lm(punitiveness_agg ~ crime_concerns_agg + process_agg + hostile_agg + emotions_agg + personality_agg, data = df_clean)

cat("Hierarchical regression results:\n")
hierarchical_results <- data.frame(
  Step       = 1:5,
  Predictors = c("Crime Concerns",
                 "+ Process Violations",
                 "+ Hostile Aggression",
                 "+ Emotions",
                 "+ Personality"),
  R2         = c(summary(step1)$r.squared, summary(step2)$r.squared,
                 summary(step3)$r.squared, summary(step4)$r.squared,
                 summary(step5)$r.squared),
  Delta_R2   = c(summary(step1)$r.squared,
                 summary(step2)$r.squared - summary(step1)$r.squared,
                 summary(step3)$r.squared - summary(step2)$r.squared,
                 summary(step4)$r.squared - summary(step3)$r.squared,
                 summary(step5)$r.squared - summary(step4)$r.squared)
) %>%
  mutate(R2       = round(R2, 3),
         Delta_R2 = round(Delta_R2, 3))

print(hierarchical_results, row.names = FALSE)

cat("\nModel comparison (ANOVA):\n")
print(anova(step1, step2, step3, step4, step5))
cat("\n")
```

# PHASE 11: SAVE ALL OUTPUT FILES

This phase writes all results to disk. The cleaned dataset is the canonical
artifact that downstream scripts (NLP pipeline, supplementary analyses,
manuscript figures) read from.

```{r phase11_save}
cat("=== SAVING ALL OUTPUT FILES ===\n\n")

# === CLEANED DATASET ===
cleaned_path <- file.path(OUT_DATA, "punishment_212_cleaned_data.csv")
write.csv(df_clean, cleaned_path, row.names = FALSE)
cat(sprintf("Saved: %s\n", cleaned_path))
cat(sprintf("       (%d rows × %d columns)\n\n", nrow(df_clean), ncol(df_clean)))

# === RELIABILITY TABLE ===
write.csv(alpha_table,
          file.path(OUT_TABLES, "reliability_alphas.csv"),
          row.names = FALSE)
cat("Saved: reliability_alphas.csv\n")

# === DESCRIPTIVES ===
write.csv(descriptives_table,
          file.path(OUT_TABLES, "descriptive_statistics.csv"),
          row.names = FALSE)
cat("Saved: descriptive_statistics.csv\n")

# === SENTENCE BY VIGNETTE ===
write.csv(as.data.frame(sentence_vs_anchor),
          file.path(OUT_TABLES, "sentence_by_vignette.csv"),
          row.names = FALSE)
cat("Saved: sentence_by_vignette.csv\n")

# === H1 RESULTS ===
write.csv(as.data.frame(fdr_table),
          file.path(OUT_TABLES, "H1_punitiveness_correlations.csv"),
          row.names = FALSE)
cat("Saved: H1_punitiveness_correlations.csv\n")

write.csv(item_cor_df,
          file.path(OUT_TABLES, "H1_item_level_correlations.csv"),
          row.names = FALSE)
cat("Saved: H1_item_level_correlations.csv\n")

# === H2 EXTENDED RESULTS ===
write.csv(extended_h2,
          file.path(OUT_TABLES, "h2_extended_steiger.csv"),
          row.names = FALSE)
cat("Saved: h2_extended_steiger.csv\n")

# === VIGNETTE STABILITY ===
write.csv(as.data.frame(vignette_cors),
          file.path(OUT_TABLES, "vignette_correlation_stability.csv"),
          row.names = FALSE)
cat("Saved: vignette_correlation_stability.csv\n")

# === DEMOGRAPHIC RESULTS ===
write.csv(as.data.frame(political_cors),
          file.path(OUT_TABLES, "political_group_correlations.csv"),
          row.names = FALSE)
cat("Saved: political_group_correlations.csv\n")

# === CFA RESULTS ===
write.csv(data.frame(Index = names(fit_indices),
                     Value = round(unname(fit_indices), 3)),
          file.path(OUT_TABLES, "cfa_fit_indices.csv"),
          row.names = FALSE)
cat("Saved: cfa_fit_indices.csv\n")

write.csv(std_solution,
          file.path(OUT_TABLES, "cfa_standardized_loadings.csv"),
          row.names = FALSE)
cat("Saved: cfa_standardized_loadings.csv\n")

write.csv(factor_cors,
          file.path(OUT_TABLES, "cfa_factor_correlations.csv"),
          row.names = FALSE)
cat("Saved: cfa_factor_correlations.csv\n")

# === REGRESSION RESULTS ===
write.csv(broom::tidy(cluster_reg),
          file.path(OUT_TABLES, "regression_clusters.csv"),
          row.names = FALSE)
cat("Saved: regression_clusters.csv\n")

write.csv(broom::tidy(construct_reg),
          file.path(OUT_TABLES, "regression_constructs.csv"),
          row.names = FALSE)
cat("Saved: regression_constructs.csv\n")

write.csv(hierarchical_results,
          file.path(OUT_TABLES, "hierarchical_regression.csv"),
          row.names = FALSE)
cat("Saved: hierarchical_regression.csv\n\n")
```

# Summary of key findings

```{r summary}
cat(paste(rep("=", 70), collapse = ""), "\n")
cat("SUMMARY OF KEY FINDINGS\n")
cat(paste(rep("=", 70), collapse = ""), "\n\n")

cat("SAMPLE:\n")
cat(sprintf("  Final N = %d (of %d collected)\n", nrow(df_clean), n_raw))
cat(sprintf("  Exclusions: %d total (%.1f%%)\n\n",
            n_raw - nrow(df_clean), 100 * (n_raw - nrow(df_clean)) / n_raw))

cat("HYPOTHESIS TESTS:\n\n")

# H1
cat("H1: Punitiveness positively correlated with all correlates\n")
cat(sprintf("  Construct-level: %d/%d (%.1f%%) positive correlations\n",
            h1_n_positive, h1_n_total,
            h1_n_positive / h1_n_total * 100))
cat(sprintf("  Significant after FDR: %d/%d\n", h1_n_sig_fdr, h1_n_total))
cat(sprintf("  SUPPORTED: %s\n\n",
            ifelse(h1_n_positive / h1_n_total > 0.9, "YES", "PARTIAL")))

# H2
cat("H2: Hostile Aggression > Crime Concerns\n")
cat(sprintf("  r(Punitiveness, Hostile)  = %.3f\n", r_hostile))
cat(sprintf("  r(Punitiveness, Crime)    = %.3f\n", r_crime))
cat(sprintf("  Difference                = %.3f\n", r_hostile - r_crime))
cat(sprintf("  Steiger's Z p-value       = %.4f\n", h2_pvalue))
cat(sprintf("  SUPPORTED: %s\n\n",
            ifelse(r_hostile > r_crime & h2_pvalue < .05,
                   "YES (Hostile Aggression > Crime Concerns)", "NO")))

# H3
cat("H3: Correlates positively intercorrelated\n")
cat(sprintf("  Positive correlations: %d/%d (%.1f%%)\n",
            h3_n_positive, h3_n_correlations,
            h3_n_positive / h3_n_correlations * 100))
cat(sprintf("  SUPPORTED: %s\n\n",
            ifelse(h3_n_positive / h3_n_correlations > 0.5, "YES", "NO")))

cat("KEY CORRELATES (Top 5 by |r|):\n")
print(as.data.frame(head(punitiveness_cors %>% select(Construct, r, sig), 5)),
      row.names = FALSE)
cat("\n")

cat("REGRESSION SUMMARY:\n")
cat(sprintf("  Total R² (all clusters):       %.3f\n", summary(cluster_reg)$r.squared))
cat(sprintf("  Hostile Aggression unique var: %.3f\n", unique_var["hostile_agg"]))
cat("\n")

cat(paste(rep("=", 70), collapse = ""), "\n")
cat("ANALYSIS COMPLETE\n")
cat(paste(rep("=", 70), collapse = ""), "\n")
```

# Session info

```{r session}
sessionInfo()
```

