---
title: "Punishment Study 2.1.2 — NLP Integration"
subtitle: "R-side analyses on text features produced by the Colab NLP pipeline"
author: "David G. Kamper"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    number_sections: false
  pdf_document:
    toc: true
---

```{r knitr_setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 7
)
```

# Overview

This script performs **all R-side NLP integration analyses** for Study 2 of
the Punishment 2.1.2 manuscript. It is the bridge between the upstream Colab
NLP pipeline (Notebooks 1–4, Python) and the manuscript's Study 2 results
section. Every number, every figure, every table that appears in Study 2
should ultimately be traceable to this script.

## What this script does NOT do

It does **not** re-run the NLP pipeline. The text features it consumes —
VADER sentiment, Empath lexicon counts, dictionary justification scores,
Claude/DeBERTa multi-label classifications, BGE/voyage/mpnet prototype
similarities, BERTopic clusters, UMAP coordinates — were all produced by
the four Colab notebooks and exported in `punishment_212_nlp_features.csv`.
That CSV is treated here as a fixed input.

## The central manuscript story it operationalizes

Two studies test whether punitive attitudes are organized around prosocial
correlates (crime concerns) or hostile correlates (revenge, hatred, etc.).
Study 1 (handled in `01_main_analysis.Rmd`) establishes that hostile
aggression dramatically outpredicts crime concerns. Study 2 (this script)
asks: when participants explain *why* they recommended a particular
sentence, does their language reveal the hostile substrate, or does it
mask it?

The empirical answer this script produces is a **method-class divergence**:

  - At the **surface category** level, all five classification methods
    (Dictionary, DeBERTa, Claude, BGE, voyage) read 70–96% of responses
    as prosocially framed. This is the "prosocial surface."
  - At the **aggregate prosocial–dark gap** level computed from
    embedding similarities (BGE, voyage, mpnet), the gap is essentially
    **uncorrelated** with hostile aggression (r ≈ −0.04 to −0.08).
    The cultural-default thesis holds at this level.
  - At the **classifier prosocial–dark gap** level (Claude ML, DeBERTa
    ML), and especially at the **theme-specific** level (BGE
    rehabilitation similarity), gaps **do** correlate with hostile
    aggression (r ≈ −0.17 to −0.37 for the strongest cells). High-
    hostile individuals select different prosocial themes, even
    while using the same overall amount of prosocial language.

The synthesis is that the prosocial framing is a *culturally shared
rhetorical script* whose surface vocabulary is invariant across
individuals, but whose internal *theme selection* tracks individual
hostility — particularly the suppression of rehabilitation language.

## Phases

  1. **Phase 0** — Setup and package loading
  2. **Phase 1** — Data loading from upstream pipeline outputs
  3. **Phase 2** — Integrity checks (N, presence of all needed columns)
  4. **Phase 3** — Surface prosocial dominance (top-category distributions
     per method, 5-way convergence)
  5. **Phase 4** — Facade correlation analysis (text features ×
     psychological measures)
  6. **Phase 5** — Mismatch multiverse (three competing hypotheses
     across six gap operationalizations)
  7. **Phase 6** — Theme-level analysis (rehabilitation, suffering, revenge
     similarity correlated with hostile aggression and other measures)
  8. **Phase 7** — BERTopic theme distributions (by hostile aggression
     tertile, by political group)
  9. **Phase 8** — Sentencing prediction (hierarchical regression with
     `punitiveness_8item`, NOT `punitiveness_agg`, to avoid circularity)
  10. **Phase 9** — Punitiveness prediction (hierarchical regression)
  11. **Phase 10** — Vignette stability (within-vignette replication)
  12. **Phase 11** — Save all output files

## Critical knowledge: the circularity issue

`punitiveness_agg` is constructed in `01_main_analysis.Rmd` as the mean of
the standardized 8-item attitudinal composite *and* `Sentence_z`. Their
correlation is r = 0.82 because `Sentence_z` is *inside*
`punitiveness_agg`. Predicting `Sentence_z` from `punitiveness_agg` is
therefore circular and inflates Step-1 R² spuriously.

**This script uses `punitiveness_8item` everywhere as the predictor**, not
`punitiveness_agg`. The 8-item attitudinal composite has r = 0.34 with
`Sentence_z` — properly independent.

(`punitiveness_agg` remains valid as the *target* of facade-correlation
description, where we are not modeling it together with `Sentence_z`.)

## Naming convention for "prevention" (per Dan Comment 518)

The new pipeline uses **incapacitation** consistently, with no separate
"prevention" category. Where Claude/DeBERTa multi-label labels include the
phrase `deterrence_and_prevention`, "prevention" refers to general crime
prevention rather than physical incapacitation, and we read these labels
as **deterrence** for manuscript-reporting purposes. The dictionary and
embedding methods use clean labels (deterrence, incapacitation,
rehabilitation, retribution, norm expression on the prosocial side; revenge,
suffering, degradation, exclusion, victim focus on the dark side).

## Output organization

All outputs are written to NLP-specific subdirectories of the existing
output folders, so they sit alongside the Phase-1 outputs without
collision.

  - `output/tables/nlp/` — CSV tables
  - `output/figures/nlp/` — PNG figures

# PHASE 0: SETUP AND PACKAGE LOADING

```{r phase0_paths}
# === PATH CONFIGURATION ===
# Inputs are expected to be present from prior pipeline stages:
#   - punishment_212_nlp_features.csv (496 × 411) from Notebook 4
#   - data/cleaned/punishment_212_cleaned_data.csv (496 × 145)
#       from 01_main_analysis.Rmd
#
# We do NOT need both — the NLP features CSV already contains the cleaned
# baseline columns (composites, demographics, vignette assignment). It is
# treated as the authoritative input. The cleaned data CSV from the main
# analysis is loaded only for cross-verification of N and key composites.

# Input paths
NLP_FEATURES_PATH  <- "data/nlp/punishment_212_nlp_features.csv"
CLEANED_DATA_PATH  <- "data/cleaned/punishment_212_cleaned_data.csv"

# Output directories — NLP-specific subdirectories
OUT_TABLES_NLP  <- "output/tables/nlp"
OUT_FIGURES_NLP <- "output/figures/nlp"

# Create output directories if needed
for (dir in c(OUT_TABLES_NLP, OUT_FIGURES_NLP)) {
  if (!dir.exists(dir)) dir.create(dir, recursive = TRUE)
}

cat("Output directories ready:\n")
cat("  Tables:  ", normalizePath(OUT_TABLES_NLP),  "\n")
cat("  Figures: ", normalizePath(OUT_FIGURES_NLP), "\n")
```

```{r phase0_packages}
# === PACKAGE LOADING ===
# All packages used downstream are listed here, with annotated purpose.

required_packages <- c(
  "tidyverse",      # Core data manipulation, ggplot2 figures
  "psych",          # Cronbach's alpha (sanity-checking composites if needed)
  "corrplot",       # Correlation heatmaps
  "Hmisc",          # rcorr() for correlation matrices with p-values
  "cocor",          # Steiger's Z-test (for any dependent-correlation comparisons)
  "lme4",           # Mixed-effects models (vignette random intercept)
  "lmerTest",       # P-values for mixed-effects models
  "car",            # Type-III ANOVA, regression diagnostics
  "effectsize",     # Cohen's d, eta-squared
  "RColorBrewer",   # Color palettes for figures
  "knitr",          # Tables for knitted output
  "broom",          # tidy() for model outputs
  "scales",         # Percentage formatting
  "irr"             # Cohen's kappa for cross-method convergence
)

new_packages <- required_packages[!(required_packages %in% installed.packages()[, "Package"])]
if (length(new_packages)) install.packages(new_packages)

invisible(lapply(required_packages, library, character.only = TRUE))

# Numeric display options — match Phase-1 settings for consistency
options(scipen = 999)
options(digits = 4)
```

```{r phase0_helpers}
# === HELPER FUNCTIONS ===
# These mirror (and where needed extend) the helpers defined in
# 01_main_analysis.Rmd. They are redefined here so this script is
# self-contained and can be run independently.

#' Format a p-value for display
format_p <- function(p) {
  if (is.na(p)) return("NA")
  if (p < .001) return("< .001")
  return(sprintf("= %.3f", p))
}

#' Convert p-value to significance stars
sig_stars <- function(p) {
  case_when(
    is.na(p) ~ "",
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ ""
  )
}

#' Compute correlation with 95% CI between two variables in a data frame
get_cor_ci <- function(x, y, data) {
  test <- cor.test(data[[x]], data[[y]])
  c(r        = unname(test$estimate),
    ci_lower = test$conf.int[1],
    ci_upper = test$conf.int[2],
    p        = test$p.value)
}

#' Pairwise raw agreement (proportion of cases where two columns are equal)
#' Used for cross-method convergence on prosocial/dark broad classification.
#'
#' Treats both NA and empty strings as missing — important because the
#' Dictionary classifier returns "" for responses with no dictionary
#' matches. Without this, the empty-string Dictionary rows count as
#' disagreements with whatever the other method says, biasing pairwise
#' agreement downward (e.g., Dict × Claude shows 0.802 instead of the
#' correct 0.945).
raw_agreement <- function(x, y) {
  ok <- !is.na(x) & !is.na(y) &
        as.character(x) != "" & as.character(y) != ""
  if (sum(ok) == 0) return(NA_real_)
  mean(x[ok] == y[ok])
}

#' Cohen's kappa for two factors
#' Treats NA and empty strings as missing (see raw_agreement note).
cohens_kappa <- function(x, y) {
  ok <- !is.na(x) & !is.na(y) &
        as.character(x) != "" & as.character(y) != ""
  if (sum(ok) < 2) return(NA_real_)
  k <- irr::kappa2(data.frame(x = x[ok], y = y[ok]))
  unname(k$value)
}

#' One-step facade correlation: a single (text feature × psych measure) cell.
#' Returns r, p, n, and a significance flag.
facade_cell <- function(data, text_var, psych_var) {
  if (!(text_var %in% names(data)) || !(psych_var %in% names(data))) {
    return(data.frame(text = text_var, psych = psych_var,
                      r = NA, p = NA, n = 0, sig = ""))
  }
  ok <- complete.cases(data[[text_var]], data[[psych_var]])
  if (sum(ok) < 3) {
    return(data.frame(text = text_var, psych = psych_var,
                      r = NA, p = NA, n = sum(ok), sig = ""))
  }
  test <- cor.test(data[[text_var]][ok], data[[psych_var]][ok])
  data.frame(
    text = text_var,
    psych = psych_var,
    r    = round(unname(test$estimate), 3),
    p    = test$p.value,
    n    = sum(ok),
    sig  = sig_stars(test$p.value)
  )
}

cat("Helper functions loaded:\n")
cat("  format_p()       — p-value formatting\n")
cat("  sig_stars()      — p → stars\n")
cat("  get_cor_ci()     — correlation with 95% CI\n")
cat("  raw_agreement()  — pairwise agreement proportion\n")
cat("  cohens_kappa()   — Cohen's kappa for two factors\n")
cat("  facade_cell()    — single text × psych correlation cell\n")
```

# PHASE 1: DATA LOADING FROM UPSTREAM PIPELINE OUTPUTS

The NLP features CSV is the master input. It already contains all 145
cleaned-data columns from `01_main_analysis.Rmd` plus 266 NLP features
derived in Notebooks 1–4 (96 + 411 = a few duplicated columns drop on
merge in the Python pipeline; the final CSV is 411 columns wide). The
cleaned-data CSV is loaded only for cross-verification.

## 1.1 Load the NLP features CSV (primary input)

```{r phase1_load_nlp}
# Read the master NLP features CSV. This is the single authoritative
# input — every downstream analysis pulls columns from here.
nlp <- read.csv(NLP_FEATURES_PATH, stringsAsFactors = FALSE)

cat(sprintf("NLP features loaded: %d rows × %d columns\n",
            nrow(nlp), ncol(nlp)))
cat(sprintf("File: %s\n", NLP_FEATURES_PATH))
```

## 1.2 Load the cleaned data CSV (cross-verification only)

```{r phase1_load_cleaned}
# We load the cleaned data file only to verify that the NLP features CSV
# correctly inherited the cleaned baseline. Any mismatch in N or in key
# composite means would indicate something went wrong upstream.

if (file.exists(CLEANED_DATA_PATH)) {
  cleaned <- read.csv(CLEANED_DATA_PATH, stringsAsFactors = FALSE)
  cat(sprintf("Cleaned data loaded: %d rows × %d columns\n",
              nrow(cleaned), ncol(cleaned)))
} else {
  cat("NOTE: cleaned data file not found at expected path.\n")
  cat("Continuing with NLP features CSV only (it contains the cleaned baseline).\n")
  cleaned <- NULL
}
```

# PHASE 2: INTEGRITY CHECKS

Before any analysis, confirm that the NLP features CSV has the structure
we expect: correct N, presence of all psychological composites from the
main analysis, presence of all NLP feature groups from the four Colab
notebooks, no unexpected missingness on the variables that drive the
manuscript's Study 2 numbers.

## 2.1 Confirm N

```{r phase2_n_check}
expected_n <- 496
actual_n   <- nrow(nlp)

cat(sprintf("Expected N: %d\n", expected_n))
cat(sprintf("Actual N:   %d\n", actual_n))
cat(sprintf("Match:      %s\n",
            ifelse(actual_n == expected_n, "YES ✓", "NO — INVESTIGATE")))

# Also confirm against cleaned data if available
if (!is.null(cleaned)) {
  cat(sprintf("\nCleaned data N: %d (should equal NLP N): %s\n",
              nrow(cleaned),
              ifelse(nrow(cleaned) == nrow(nlp), "YES ✓", "MISMATCH")))
}
```

## 2.2 Confirm presence of psychological composites from Phase 1

```{r phase2_psych_composites}
# These composites were built in 01_main_analysis.Rmd. They must all be
# present in the NLP CSV — if any are missing, the export from the
# Python pipeline failed to inherit them and we need to investigate.

required_psych <- c(
  # Punitiveness DV (both forms — circular and non-circular)
  "punitiveness_agg", "punitiveness_8item", "punitiveness_8item_z", "Sentence_z",
  # Cluster aggregates
  "crime_concerns_agg", "process_agg", "emotions_agg",
  "hostile_agg", "personality_agg",
  # Construct-level composites (used in facade matrix)
  "hatred_comp", "anger_comp",
  "exclusion_comp", "degradation_comp", "suffering_comp",
  "prisonvi_comp", "harsh_comp", "revenge_comp",
  "rwa_comp", "sdo_comp", "venge_comp", "vprone_comp",
  "raceresent_comp", "bloodsports_comp",
  "crime_rates_comp", "fear_comp",
  "dueprocess_comp", "uncertain_comp",
  # Demographics & moderators
  "age_1", "gender", "education", "race", "politid", "political_group",
  # Vignette
  "vignette"
)

missing_psych <- setdiff(required_psych, names(nlp))

cat("=== PSYCHOLOGICAL COMPOSITE CHECK ===\n\n")
cat(sprintf("Required: %d\n", length(required_psych)))
cat(sprintf("Present:  %d\n", length(required_psych) - length(missing_psych)))
cat(sprintf("Missing:  %d\n\n", length(missing_psych)))

if (length(missing_psych) > 0) {
  cat("MISSING COLUMNS (investigate before continuing):\n")
  for (m in missing_psych) cat("  -", m, "\n")
  cat("\n")
} else {
  cat("All required composites present ✓\n\n")
}
```

## 2.3 Confirm presence of NLP feature groups

```{r phase2_nlp_features}
# Each row in this metadata table documents one NLP feature group:
#   - prefix:        column-name pattern that identifies the group
#   - source:        which Colab notebook produced these columns
#   - rationale:     what these features measure
#   - expected_min:  minimum number of columns we expect (sanity floor)
#
# This serves as the auditable manifest for what the NLP pipeline is
# supposed to deliver to the R side.

nlp_feature_spec <- tibble::tribble(
  ~group,          ~prefix,    ~source,                     ~rationale,                                                                ~expected_min,

  "VADER",         "vader",    "Notebook 1 (preprocessing)", "Sentiment intensity (compound, neg, neu, pos) — surface affect proxy",   12,
  "Empath",        "empath_",  "Notebook 1 (preprocessing)", "Lexicon-based topic counts (violence, healing, trust, etc.)",           20,
  "Dictionary",    "just_",    "Notebook 1 (preprocessing)", "Custom justification dictionary (deterrence, rehab, revenge, etc.)",    13,
  "Claude",        "claude_",  "Notebook 2 (classification)", "Anthropic Claude Haiku multi-label + forced-choice classifications",   15,
  "DeBERTa",       "deberta_", "Notebook 2 (classification)", "DeBERTa-v3 zero-shot multi-label + forced-choice classifications",     20,
  "mpnet",         "mpnet_",   "Notebook 3 (embeddings)",    "MPNet (2021 baseline) prototype cosine similarities",                   40,
  "BGE",           "bge_",     "Notebook 3 (embeddings)",    "BGE-large (2024) prototype cosine similarities — primary workhorse",    40,
  "voyage",        "voyage_",  "Notebook 3 (embeddings)",    "voyage-4-large (2026) prototype cosine similarities — cross-validation", 40,
  "BERTopic/UMAP", "bertopic", "Notebook 4 (topics + export)", "Topic cluster IDs and 2-D UMAP coordinates",                          2,
  "Method tops",   "method_",  "Notebook 4 (topics + export)", "Top-category labels per method (used in convergence analyses)",       5
)

cat("=== NLP FEATURE GROUP CHECK ===\n\n")

feature_check <- nlp_feature_spec %>%
  rowwise() %>%
  mutate(
    n_present = sum(grepl(paste0("^", prefix), names(nlp))),
    status    = ifelse(n_present >= expected_min, "✓", "✗")
  ) %>%
  ungroup()

print(as.data.frame(feature_check %>% select(group, prefix, n_present, expected_min, status)),
      row.names = FALSE)

cat(sprintf("\nTotal columns matching feature-group prefixes: %d\n",
            sum(feature_check$n_present)))
cat(sprintf("Total columns in NLP CSV: %d\n", ncol(nlp)))
```

## 2.4 Confirm response-text columns are present and populated

```{r phase2_text_columns}
# The actual open-ended responses must be present so we can spot-check
# coverage and word-count distributions. If text columns are missing or
# heavily empty, all downstream NLP features are suspect.

text_cols <- c("Sentence_OpenEnd1", "Sentence_OpenEnd2",
               "text1_clean", "text2_clean", "text_combined",
               "text_combined_wordcount")

text_check <- data.frame(
  Column      = text_cols,
  Present     = text_cols %in% names(nlp),
  N_NonEmpty  = sapply(text_cols, function(c) {
    if (!(c %in% names(nlp))) return(NA)
    sum(!is.na(nlp[[c]]) & nchar(as.character(nlp[[c]])) > 0)
  })
)

cat("=== TEXT COLUMN CHECK ===\n\n")
print(text_check, row.names = FALSE)

cat(sprintf("\nWord count summary (text_combined_wordcount):\n"))
print(summary(nlp$text_combined_wordcount))
cat("\n")
```

## 2.5 Spot-check missingness on key analysis variables

```{r phase2_missingness}
# For variables that drive the manuscript's headline numbers, any
# unexpected missingness needs to surface here, not buried in a
# regression error message later.

key_vars <- c(
  "Sentence_z", "punitiveness_agg", "punitiveness_8item",
  "hostile_agg", "crime_concerns_agg",
  "bge_orig_rehabilitation", "bge_orig_b3a_pro_minus_dark",
  "claude_ml_prosocial_minus_dark", "deberta_ml_prosocial_minus_dark",
  "bertopic_topic", "vader_compound"
)

missing_summary <- nlp %>%
  select(any_of(key_vars)) %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "N_Missing") %>%
  mutate(Pct_Missing = round(100 * N_Missing / nrow(nlp), 1))

cat("=== MISSINGNESS ON KEY ANALYSIS VARIABLES ===\n\n")
print(as.data.frame(missing_summary), row.names = FALSE)
cat("\n")
```

## 2.6 Cross-verify against cleaned data (if loaded)

```{r phase2_cross_verify}
# If we have both the cleaned data and the NLP features CSV, cross-check
# that core composite means match exactly. If they don't, the Python
# pipeline diverged from the R baseline in a way that needs investigation.

if (!is.null(cleaned)) {
  check_vars <- c("punitiveness_agg", "punitiveness_8item", "hostile_agg",
                  "crime_concerns_agg", "Sentence_z")

  cat("=== CROSS-VERIFICATION: NLP CSV vs cleaned data CSV ===\n\n")
  cat(sprintf("%-22s  %-12s  %-12s  %-10s\n",
              "Variable", "NLP_mean", "Cleaned_mean", "Match"))
  cat(strrep("-", 60), "\n")

  for (v in check_vars) {
    if (v %in% names(nlp) && v %in% names(cleaned)) {
      m_nlp     <- mean(nlp[[v]],     na.rm = TRUE)
      m_cleaned <- mean(cleaned[[v]], na.rm = TRUE)
      match_ok  <- abs(m_nlp - m_cleaned) < 1e-6

      cat(sprintf("%-22s  %-12.4f  %-12.4f  %-10s\n",
                  v, m_nlp, m_cleaned,
                  ifelse(match_ok, "✓", "✗ MISMATCH")))
    }
  }
  cat("\n")
} else {
  cat("Cleaned data CSV not loaded — skipping cross-verification.\n\n")
}
```


# PHASE 3: SURFACE PROSOCIAL DOMINANCE

The first manuscript-relevant question is: **at the surface category
level, what proportion of responses do classifiers read as prosocial vs.
dark?** If the prosocial framing is universal (as the cultural-default
thesis predicts), then *all* methods should converge on a high prosocial
percentage. This phase produces the descriptive backbone for "The
Prosocial Surface" subsection of Study 2.

## 3.1 Top-category distributions per method

```{r phase3_top_category_distributions}
# Each method has produced a top-category label per response. We tabulate
# them all here. Note that label vocabularies differ across methods —
# Dictionary uses our 10 categories, Claude/DeBERTa use the 8 multi-label
# candidate categories from Notebook 2, BGE/voyage use the embedding
# prototype labels from Notebook 3.

cat("=== TOP-CATEGORY DISTRIBUTIONS PER METHOD ===\n\n")

method_top_cols <- c(
  "Dictionary"   = "method_dict_top",
  "DeBERTa"      = "method_deberta_top",
  "Claude"       = "method_claude_top",
  "BGE"          = "method_bge_top",
  "voyage"       = "method_voyage_top"
)

for (method in names(method_top_cols)) {
  col <- method_top_cols[[method]]
  if (!(col %in% names(nlp))) {
    cat(sprintf("[%s] column not found, skipping\n\n", method))
    next
  }

  tab <- sort(table(nlp[[col]], useNA = "ifany"), decreasing = TRUE)
  pct <- round(100 * tab / sum(tab), 1)

  cat(sprintf("--- %s (n = %d) ---\n", method, sum(tab)))
  for (i in seq_along(tab)) {
    cat(sprintf("  %-32s %4d (%.1f%%)\n",
                names(tab)[i], tab[i], pct[i]))
  }
  cat("\n")
}
```

## 3.2 Broad prosocial vs. dark per method

The substantive manuscript number is the % prosocial vs. % dark — that
is, the binary collapse of each method's top category onto our broad
prosocial/dark schema. The CSV already includes precomputed
`*_top_broad` columns from the Python pipeline; we re-tabulate them
here in clean R-native form for export.

```{r phase3_prosocial_dark_broad}
broad_cols <- c(
  "Dictionary"   = "dict_top_broad",
  "DeBERTa"      = "deberta_top_broad",
  "Claude"       = "claude_top_broad",
  "BGE"          = "bge_top_broad",
  "voyage"       = "voyage_top_broad"
)

# Build the manuscript-ready summary table
broad_summary <- data.frame(
  Method = character(),
  N      = integer(),
  Pct_Prosocial  = numeric(),
  Pct_Dark       = numeric(),
  N_Unclassified = integer(),
  stringsAsFactors = FALSE
)

cat("=== PROSOCIAL VS DARK (TOP-CATEGORY → BROAD) ===\n\n")

for (method in names(broad_cols)) {
  col <- broad_cols[[method]]
  if (!(col %in% names(nlp))) {
    cat(sprintf("[%s] column not found, skipping\n", method))
    next
  }

  # IMPORTANT: read.csv keeps "" (empty strings) as literal character values,
  # not as NA. The Dictionary method produces "" for responses where no
  # dictionary term matched at all (75 such rows in this dataset). We treat
  # empty strings as missing here so they don't dilute the prosocial-vs-dark
  # denominator. This matters: without this filter, Dictionary's prosocial
  # rate appears as 406/496 = 81.9%, conflating "Dictionary said prosocial"
  # with "Dictionary said nothing." The cleanly-comparable rate, given that
  # the four other methods classify every response, is 406/421 = 96.4%.
  vals <- nlp[[col]]
  vals <- vals[!is.na(vals) & vals != ""]
  tab <- table(vals)
  n_total <- sum(tab)

  pct_pro  <- if ("prosocial" %in% names(tab)) 100 * tab[["prosocial"]] / n_total else 0
  pct_dark <- if ("dark"      %in% names(tab)) 100 * tab[["dark"]]      / n_total else 0
  n_unclassified <- sum(is.na(nlp[[col]]) | nlp[[col]] == "")

  cat(sprintf("  %-12s n=%3d   prosocial = %5.1f%%   dark = %5.1f%%   (unclassified: %d)\n",
              method, n_total, pct_pro, pct_dark, n_unclassified))

  broad_summary <- rbind(broad_summary, data.frame(
    Method = method,
    N      = n_total,
    Pct_Prosocial  = round(pct_pro,  1),
    Pct_Dark       = round(pct_dark, 1),
    N_Unclassified = n_unclassified
  ))
}

cat(sprintf("\n*** SECTION SUMMARY (Phase 3.2): "))
cat(sprintf("Range of prosocial = %.1f%% to %.1f%% across methods. ",
            min(broad_summary$Pct_Prosocial),
            max(broad_summary$Pct_Prosocial)))
cat("All methods read majority prosocial. ***\n\n")
```

## 3.3 Cross-method convergence (raw agreement and Cohen's κ)

A critical reviewer question is whether the 5 methods are agreeing or
just all recovering the same base-rate skew. We compute pairwise raw
agreement and Cohen's κ for every method-pair on the broad
prosocial/dark classification.

The κ values will be much smaller than raw-agreement values because of
the "kappa paradox" under heavy class imbalance (when the true rate is
~95% prosocial, two methods that are both ~95% accurate will have raw
agreement ≈ 0.95 but κ ≈ 0). We report both and discuss this in the
manuscript supplementary.

```{r phase3_convergence}
methods_for_conv <- names(broad_cols)
n_methods <- length(methods_for_conv)

# All unordered pairs
pairs <- t(combn(methods_for_conv, 2))

convergence <- data.frame(
  Method_1      = pairs[, 1],
  Method_2      = pairs[, 2],
  N             = NA_integer_,
  Raw_Agreement = NA_real_,
  Kappa         = NA_real_,
  stringsAsFactors = FALSE
)

for (i in seq_len(nrow(pairs))) {
  m1 <- broad_cols[[pairs[i, 1]]]
  m2 <- broad_cols[[pairs[i, 2]]]

  if (!(m1 %in% names(nlp)) || !(m2 %in% names(nlp))) next

  # Effective N: rows where BOTH methods produced a classification
  # (i.e., neither NA nor empty string). For non-Dictionary pairs this
  # equals 496; for Dictionary pairs it equals 421.
  ok <- !is.na(nlp[[m1]]) & !is.na(nlp[[m2]]) &
        as.character(nlp[[m1]]) != "" & as.character(nlp[[m2]]) != ""
  convergence$N[i]             <- sum(ok)
  convergence$Raw_Agreement[i] <- round(raw_agreement(nlp[[m1]], nlp[[m2]]), 3)
  convergence$Kappa[i]         <- round(cohens_kappa(nlp[[m1]], nlp[[m2]]), 3)
}

cat("=== 5-WAY CROSS-METHOD CONVERGENCE ===\n\n")
print(convergence, row.names = FALSE)

cat(sprintf("\nRaw agreement range: %.3f to %.3f\n",
            min(convergence$Raw_Agreement, na.rm = TRUE),
            max(convergence$Raw_Agreement, na.rm = TRUE)))
cat(sprintf("Cohen's κ range:     %.3f to %.3f\n",
            min(convergence$Kappa, na.rm = TRUE),
            max(convergence$Kappa, na.rm = TRUE)))

cat("\nINTERPRETATION:\n")
cat("  - High raw agreement + low κ = base-rate skew (κ paradox).\n")
cat("  - All five methods converge on majority prosocial; classifier-classifier\n")
cat("    pairs agree more on the minority dark class (e.g., BGE × voyage κ = 0.31).\n")
cat("  - Report raw agreement primarily; κ in supplementary with paradox footnote.\n\n")
```

## 3.4 Save Phase 3 outputs

```{r phase3_save}
write.csv(broad_summary,
          file.path(OUT_TABLES_NLP, "method_top_category_summary.csv"),
          row.names = FALSE)
cat("Saved: method_top_category_summary.csv\n")

write.csv(convergence,
          file.path(OUT_TABLES_NLP, "cross_method_convergence_5way.csv"),
          row.names = FALSE)
cat("Saved: cross_method_convergence_5way.csv\n\n")
```

# PHASE 4: FACADE CORRELATION ANALYSIS

This is the central descriptive table for Study 2's "Three Competing
Explanations" subsection. We correlate every text feature (rows) with
every psychological measure (columns) and visualize the result as a
heatmap. The cultural-default thesis predicts most cells should be near
zero. The individual-facade thesis predicts negative cells (high-hostile
→ less prosocial language). Theme-selection effects show up as
*selectively* large negative cells (e.g., rehab similarity), while other
prosocial themes look flat.

## 4.1 Define the rows (text features) and columns (psychological measures)

```{r phase4_specs}
# === TEXT FEATURES (rows of facade matrix) ===
# We pick a curated set that spans (a) embedding-based prosocial-dark gaps
# (one per model), (b) classifier-based prosocial-dark gaps, (c) individual
# theme-similarity columns from the workhorse embedding model (BGE), and
# (d) low-level surface features (sentiment, word count).

text_features <- tibble::tribble(
  ~text_var,                          ~text_label,                ~feature_type,
  "bge_orig_b3a_pro_minus_dark",      "BGE pro−dark gap",         "embedding gap",
  "voyage_orig_b3a_pro_minus_dark",   "voyage pro−dark gap",      "embedding gap",
  "mpnet_orig_b3a_pro_minus_dark",    "mpnet pro−dark gap",       "embedding gap",
  "claude_ml_prosocial_minus_dark",   "Claude pro−dark (ML)",     "classifier gap",
  "deberta_ml_prosocial_minus_dark",  "DeBERTa pro−dark (ML)",    "classifier gap",
  "just_prosocial_minus_dark",        "Dictionary pro−dark",      "dictionary gap",
  "vader_compound",                   "VADER compound",           "sentiment",
  "text_combined_wordcount",          "Word count",               "surface",
  "bge_orig_rehabilitation",          "BGE rehab sim",            "theme similarity",
  "bge_orig_suffering",               "BGE suffering sim",        "theme similarity",
  "bge_orig_revenge",                 "BGE revenge sim",          "theme similarity"
)

# === PSYCHOLOGICAL MEASURES (columns of facade matrix) ===
# Mix of cluster aggregates (Punitiveness, Hostile Aggression, Crime
# Concerns) and theoretically informative construct-level composites
# (Hatred, Revenge, Suffering, Exclusion, RWA, Racial Resentment).

psych_measures <- tibble::tribble(
  ~psych_var,             ~psych_label,
  "punitiveness_agg",     "Punitiveness",
  "hostile_agg",          "Hostile Aggression",
  "crime_concerns_agg",   "Crime Concerns",
  "hatred_comp",          "Hatred",
  "revenge_comp",         "Revenge",
  "suffering_comp",       "Suffering",
  "exclusion_comp",       "Exclusion",
  "rwa_comp",             "RWA",
  "raceresent_comp",      "Racial Resentment"
)

cat(sprintf("Facade matrix: %d text features × %d psych measures = %d cells\n\n",
            nrow(text_features), nrow(psych_measures),
            nrow(text_features) * nrow(psych_measures)))
```

## 4.2 Compute the full facade correlation matrix

```{r phase4_compute_matrix}
# Loop over every (text, psych) pair and compute one facade_cell()
facade_long <- expand.grid(
  text_var  = text_features$text_var,
  psych_var = psych_measures$psych_var,
  stringsAsFactors = FALSE
)

facade_results <- do.call(rbind, lapply(seq_len(nrow(facade_long)), function(i) {
  facade_cell(nlp,
              facade_long$text_var[i],
              facade_long$psych_var[i])
}))

# Attach human-readable labels
facade_results <- facade_results %>%
  left_join(text_features  %>% select(text_var,  text_label),
            by = c("text"  = "text_var")) %>%
  left_join(psych_measures %>% select(psych_var, psych_label),
            by = c("psych" = "psych_var"))

# Sort by absolute correlation
facade_results <- facade_results %>%
  arrange(desc(abs(r)))

cat("=== FACADE CORRELATION TABLE (top 20 cells by |r|) ===\n\n")
print(facade_results %>%
        select(TextFeature = text_label, PsychMeasure = psych_label,
               r, p, n, sig) %>%
        head(20),
      row.names = FALSE)

# Summary stats
n_total       <- nrow(facade_results)
n_sig         <- sum(facade_results$p < .05, na.rm = TRUE)
n_meaningful  <- sum(facade_results$p < .05 & abs(facade_results$r) >= .10, na.rm = TRUE)
n_neg_meaning <- sum(facade_results$p < .05 & facade_results$r <= -.10, na.rm = TRUE)
n_pos_meaning <- sum(facade_results$p < .05 & facade_results$r >=  .10, na.rm = TRUE)

cat(sprintf("\n*** SECTION SUMMARY (Phase 4.2):\n"))
cat(sprintf("  Cells:                     %d\n", n_total))
cat(sprintf("  Significant (p < .05):     %d (%.1f%%)\n",
            n_sig, 100 * n_sig / n_total))
cat(sprintf("  Meaningful (|r|≥.10, p<.05): %d (%.1f%%)\n",
            n_meaningful, 100 * n_meaningful / n_total))
cat(sprintf("  Negative meaningful:       %d\n", n_neg_meaning))
cat(sprintf("  Positive meaningful:       %d\n", n_pos_meaning))
cat("  → Pattern: where text features track psychological measures,\n")
cat("    they almost always do so NEGATIVELY (prosocial framing decreases\n")
cat("    with punitiveness/hostility) — consistent with theme-selection effects. ***\n\n")
```

## 4.3 Facade heatmap visualization

```{r phase4_heatmap}
# Build the heatmap with ggplot2 so we can combine r-value and significance
# stars in each cell. corrplot doesn't support both in one tile cleanly.
#
# Each cell shows:
#   - background color: gradient red (negative) → white (zero) → blue (positive)
#   - text:             the r value, followed by significance stars
#                       (* p < .05, ** p < .01, *** p < .001)
#
# Row and column order is preserved from text_features and psych_measures
# specs in section 4.1, so the visual order matches the manuscript narrative
# (embedding gaps, classifier gaps, dictionary gap, surface features,
#  theme similarities — top to bottom).

facade_plot_data <- facade_results %>%
  mutate(
    # Add stars to the r-value label
    cell_label = sprintf("%.2f%s", r, sig),
    # Force the row order: use the order of text_features
    text_label = factor(text_label, levels = rev(text_features$text_label)),
    # Force the column order: use the order of psych_measures
    psych_label = factor(psych_label, levels = psych_measures$psych_label)
  )

facade_heatmap <- ggplot(facade_plot_data,
                          aes(x = psych_label, y = text_label, fill = r)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = cell_label), size = 3, color = "black") +
  scale_fill_gradient2(
    low      = "#B2182B",     # red for negative
    mid      = "white",
    high     = "#2166AC",     # blue for positive
    midpoint = 0,
    limits   = c(-0.40, 0.40),
    name     = "Pearson r"
  ) +
  scale_x_discrete(position = "top") +
  labs(
    title    = "Facade Analysis: Text Features × Psychological Measures",
    subtitle = "Significance: * p < .05, ** p < .01, *** p < .001",
    x        = NULL,
    y        = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x       = element_text(angle = 45, hjust = 0, vjust = 0,
                                      face = "bold"),
    axis.text.y       = element_text(face = "bold"),
    panel.grid        = element_blank(),
    plot.title        = element_text(face = "bold", size = 13, hjust = 0.5),
    plot.subtitle     = element_text(size = 9, hjust = 0.5,
                                      color = "grey30"),
    legend.position   = "right",
    legend.key.height = unit(1.5, "cm")
  )

# Save to file
ggsave(filename = file.path(OUT_FIGURES_NLP, "facade_correlations_heatmap.png"),
       plot     = facade_heatmap,
       width    = 11,
       height   = 8,
       dpi      = 150,
       bg       = "white")

cat("Saved: facade_correlations_heatmap.png\n")

# Also display in the rendered HTML
print(facade_heatmap)
```

## 4.4 Save full facade table

```{r phase4_save}
write.csv(facade_results %>%
            select(TextFeature = text_label, PsychMeasure = psych_label,
                   r, p, N = n, sig),
          file.path(OUT_TABLES_NLP, "facade_correlations_full.csv"),
          row.names = FALSE)
cat("Saved: facade_correlations_full.csv\n\n")
```

# PHASE 5: MISMATCH MULTIVERSE — THREE COMPETING HYPOTHESES

This phase operationalizes the manuscript's "Three Competing
Explanations" subsection. We test, for each of six different
prosocial-dark gap operationalizations, whether hostile aggression and
crime concerns correlate with the gap, and assign a verdict among:

  - **Individual Facade**: r(hostile, gap) is significantly negative
    (high-hostile individuals use less prosocial language to mask dark
    motives).
  - **Sincerity**: r(crime, gap) is significantly positive
    (high-crime-concern individuals use more prosocial language because
    they genuinely have prosocial motives).
  - **Cultural Default**: both correlations are near zero
    (everyone uses similar amounts of prosocial language; the
    framing is a shared rhetorical script independent of psychology).

Per Dan Comment 519, these three hypotheses are described before
predictions are stated.

## 5.1 Define the six gap operationalizations

```{r phase5_gap_specs}
# Three embedding-based gaps (BGE, voyage, mpnet) and three
# classifier-based gaps (Claude ML, DeBERTa ML, Dictionary).

gap_specs <- tibble::tribble(
  ~gap_var,                            ~gap_label,
  "bge_orig_b3a_pro_minus_dark",       "BGE",
  "voyage_orig_b3a_pro_minus_dark",    "voyage",
  "mpnet_orig_b3a_pro_minus_dark",     "mpnet",
  "claude_ml_prosocial_minus_dark",    "Claude ML",
  "deberta_ml_prosocial_minus_dark",   "DeBERTa ML",
  "just_prosocial_minus_dark",         "Dictionary"
)

cat(sprintf("Testing %d gap operationalizations across 3 hypotheses each.\n\n",
            nrow(gap_specs)))
```

## 5.2 Compute the multiverse table

```{r phase5_multiverse}
# Verdict logic:
#   - If r(hostile) is sig negative AND |r(hostile)| > |r(crime)| → Individual Facade
#   - Else if r(crime) is sig positive → Sincerity
#   - Else → Cultural Default
# This implements the description in Dan-comment-519 ordering: predictions
# stated AFTER hypotheses described above.

multiverse <- data.frame(
  TextFeature = character(),
  r_hostile   = numeric(),
  p_hostile   = numeric(),
  r_crime     = numeric(),
  p_crime     = numeric(),
  Verdict     = character(),
  stringsAsFactors = FALSE
)

for (i in seq_len(nrow(gap_specs))) {
  gv <- gap_specs$gap_var[i]
  gl <- gap_specs$gap_label[i]

  if (!(gv %in% names(nlp))) next

  # r(gap, hostile_agg)
  th <- cor.test(nlp[[gv]], nlp$hostile_agg)
  rh <- unname(th$estimate); ph <- th$p.value

  # r(gap, crime_concerns_agg)
  tc <- cor.test(nlp[[gv]], nlp$crime_concerns_agg)
  rc <- unname(tc$estimate); pc <- tc$p.value

  verdict <- if (ph < .05 && rh < 0 && abs(rh) > abs(rc)) {
    "INDIVIDUAL FACADE"
  } else if (pc < .05 && rc > 0) {
    "SINCERITY"
  } else {
    "CULTURAL DEFAULT"
  }

  multiverse <- rbind(multiverse, data.frame(
    TextFeature = gl,
    r_hostile   = round(rh, 4),
    p_hostile   = round(ph, 4),
    r_crime     = round(rc, 4),
    p_crime     = round(pc, 4),
    Verdict     = verdict
  ))
}

cat("=== MULTIVERSE TEST OF MISMATCH HYPOTHESES ===\n\n")
print(multiverse, row.names = FALSE)

n_cd <- sum(multiverse$Verdict == "CULTURAL DEFAULT")
n_if <- sum(multiverse$Verdict == "INDIVIDUAL FACADE")
n_si <- sum(multiverse$Verdict == "SINCERITY")

cat(sprintf("\n*** SECTION SUMMARY (Phase 5.2):\n"))
cat(sprintf("  Cultural Default:   %d/%d  (embedding-based gaps)\n",
            n_cd, nrow(multiverse)))
cat(sprintf("  Individual Facade:  %d/%d  (classifier-based gaps)\n",
            n_if, nrow(multiverse)))
cat(sprintf("  Sincerity:          %d/%d\n", n_si, nrow(multiverse)))
cat("  → The verdict pattern reveals method-class divergence: surface\n")
cat("    prosocial-vocabulary similarity (embeddings) is independent\n")
cat("    of hostility, but classifier-detected theme selection IS\n")
cat("    associated with hostility. Both findings are robust within\n")
cat("    their respective method classes. ***\n\n")
```

## 5.3 Save multiverse table

```{r phase5_save}
write.csv(multiverse,
          file.path(OUT_TABLES_NLP, "mismatch_multiverse.csv"),
          row.names = FALSE)
cat("Saved: mismatch_multiverse.csv\n\n")
```


# PHASE 6: THEME-LEVEL ANALYSIS

The multiverse phase (5) showed that *aggregate* prosocial-dark gaps from
embeddings are uncorrelated with hostility, but classifier gaps are. This
phase asks the more fine-grained question: **which specific themes carry
the signal?** The strongest single facade cell in Phase 4 was BGE
rehabilitation similarity × Punitiveness (r ≈ −0.37). We follow that
thread here.

The hypothesis is that high-punitive / high-hostile individuals do not
suppress prosocial language uniformly — they suppress *rehabilitation*
specifically while continuing to invoke deterrence and proportional
justice. If true, this would explain why the embedding-based aggregate
gap looks flat (rehabilitation is one of seven prototypes, and other
prosocial themes pick up the slack) while the dictionary and classifier
methods pick up theme selection more cleanly.

## 6.1 BGE prototype similarities × hostile aggression and related measures

```{r phase6_bge_themes}
# All seven BGE prototype similarities from the "original" prototype set.
# Tagged with prosocial/dark side for display purposes.
bge_themes <- tibble::tribble(
  ~theme_var,                ~side,
  "bge_orig_deterrence",     "prosocial",
  "bge_orig_incapacitation", "prosocial",
  "bge_orig_rehabilitation", "prosocial",
  "bge_orig_retribution",    "prosocial",
  "bge_orig_revenge",        "dark",
  "bge_orig_suffering",      "dark",
  "bge_orig_exclusion",      "dark"
)

theme_axis <- c("hostile_agg", "punitiveness_agg", "rwa_comp",
                "raceresent_comp", "crime_concerns_agg")

theme_results <- expand.grid(
  text  = bge_themes$theme_var,
  psych = theme_axis,
  stringsAsFactors = FALSE
)

theme_results <- do.call(rbind, lapply(seq_len(nrow(theme_results)), function(i) {
  facade_cell(nlp, theme_results$text[i], theme_results$psych[i])
}))

# Attach side and short theme label
theme_results <- theme_results %>%
  left_join(bge_themes, by = c("text" = "theme_var")) %>%
  mutate(theme = gsub("^bge_orig_", "", text))

cat("=== BGE THEME SIMILARITIES × PSYCHOLOGICAL MEASURES ===\n\n")

# Pivot to wide for readable display
theme_wide <- theme_results %>%
  select(theme, side, psych, r) %>%
  pivot_wider(names_from = psych, values_from = r) %>%
  arrange(side, theme)

print(as.data.frame(theme_wide), row.names = FALSE)

cat("\nINTERPRETATION:\n")
cat("  - The rehabilitation row is consistently the most negative across\n")
cat("    psychological measures: it is selectively suppressed by punitive\n")
cat("    and high-hostile individuals.\n")
cat("  - Other prosocial themes (deterrence, incapacitation, retribution)\n")
cat("    show much weaker or null correlations.\n")
cat("  - Dark themes (revenge, suffering, exclusion) do NOT show positive\n")
cat("    correlations with hostility — high-hostile individuals are not\n")
cat("    expressing more dark language; they are suppressing rehabilitation\n")
cat("    language while leaving the dark-theme similarities essentially\n")
cat("    flat. This asymmetry is itself a key manuscript finding: theme\n")
cat("    selection in this domain is about what high-hostile individuals\n")
cat("    AVOID saying, not what they openly say.\n\n")
```

## 6.2 Cross-model verification: do voyage and mpnet show the same theme pattern?

```{r phase6_cross_model}
# If the rehabilitation-suppression pattern is real (not BGE-specific),
# voyage and mpnet should produce the same qualitative pattern.

cross_model_themes <- expand.grid(
  model     = c("bge", "voyage", "mpnet"),
  theme     = c("deterrence", "incapacitation", "rehabilitation",
                "retribution", "revenge", "suffering", "exclusion"),
  stringsAsFactors = FALSE
)
cross_model_themes$col <- paste0(cross_model_themes$model, "_orig_",
                                  cross_model_themes$theme)

# r with hostile_agg
cross_model_themes$r_hostile <- sapply(cross_model_themes$col, function(c) {
  if (!(c %in% names(nlp))) return(NA_real_)
  cor(nlp[[c]], nlp$hostile_agg, use = "pairwise.complete.obs")
})

# r with punitiveness_agg
cross_model_themes$r_punitive <- sapply(cross_model_themes$col, function(c) {
  if (!(c %in% names(nlp))) return(NA_real_)
  cor(nlp[[c]], nlp$punitiveness_agg, use = "pairwise.complete.obs")
})

cat("=== THEME-SIMILARITY × HOSTILE/PUNITIVE ACROSS THREE MODELS ===\n\n")

# Pivot to compare across models
cross_wide <- cross_model_themes %>%
  select(theme, model, r_hostile) %>%
  pivot_wider(names_from = model, values_from = r_hostile,
              names_prefix = "r_hostile_") %>%
  rename(BGE_hostile = r_hostile_bge, voyage_hostile = r_hostile_voyage, mpnet_hostile = r_hostile_mpnet)

cross_wide_pun <- cross_model_themes %>%
  select(theme, model, r_punitive) %>%
  pivot_wider(names_from = model, values_from = r_punitive,
              names_prefix = "r_punitive_") %>%
  rename(BGE_punitive = r_punitive_bge, voyage_punitive = r_punitive_voyage, mpnet_punitive = r_punitive_mpnet)

cross_combined <- cross_wide %>% left_join(cross_wide_pun, by = "theme") %>%
  mutate(across(where(is.numeric), ~ round(., 3)))

print(as.data.frame(cross_combined), row.names = FALSE)

cat("\n*** SECTION SUMMARY (Phase 6.2):\n")
cat("  - Rehabilitation-suppression pattern replicates across all three\n")
cat("    embedding models (BGE, voyage, mpnet) for both hostility AND\n")
cat("    punitiveness as the criterion.\n")
cat("  - This rules out a single-model artifact: the manuscript story\n")
cat("    'high-hostile individuals selectively suppress rehabilitation\n")
cat("    language' is robust to model choice. ***\n\n")
```

## 6.3 Save Phase 6 outputs

```{r phase6_save}
write.csv(theme_results %>%
            select(theme, side, psych, r, p, n) %>%
            mutate(across(where(is.numeric), ~ round(., 3))),
          file.path(OUT_TABLES_NLP, "bge_theme_correlations.csv"),
          row.names = FALSE)
cat("Saved: bge_theme_correlations.csv\n")

write.csv(cross_combined,
          file.path(OUT_TABLES_NLP, "theme_correlations_cross_model.csv"),
          row.names = FALSE)
cat("Saved: theme_correlations_cross_model.csv\n\n")
```

# PHASE 7: BERTOPIC THEME DISTRIBUTIONS

Notebook 4 produced 7 BERTopic clusters (plus an outlier cluster −1) from
BGE embeddings of the open-ended responses. Topics include
incapacitation/exclusion themes, vignette-specific factual content, moral
desert framing, sentence anchoring, victim focus, and rehabilitation
themes. This phase asks: **does the proportion of responses in each
topic differ by hostile-aggression tertile or by political group?**

If the cultural-default thesis holds at the topic level, distributions
should look similar across tertiles. If theme selection differs (as
Phase 6 implies for rehabilitation specifically), we should see fewer
high-hostile responses in rehabilitation-themed topics.

## 7.1 BERTopic distribution overall

```{r phase7_distribution}
cat("=== BERTOPIC TOPIC DISTRIBUTION ===\n\n")

bert_tab <- table(nlp$bertopic_topic, useNA = "ifany")
bert_pct <- round(100 * bert_tab / sum(bert_tab), 1)

bert_summary <- data.frame(
  Topic = names(bert_tab),
  N     = as.integer(bert_tab),
  Pct   = as.numeric(bert_pct)
)
print(bert_summary, row.names = FALSE)

n_outlier <- sum(nlp$bertopic_topic == -1, na.rm = TRUE)
cat(sprintf("\nOutliers (Topic -1): %d (%.1f%%)\n",
            n_outlier, 100 * n_outlier / nrow(nlp)))
cat("Note: HDBSCAN labels responses as outliers when they cannot be\n")
cat("confidently assigned to any cluster. ~28% outlier rate is typical\n")
cat("for short opinion text and is reported transparently.\n\n")
```

## 7.2 Topic distribution by hostile-aggression tertile

```{r phase7_by_tertile}
# The CSV already contains a hostile_tertile column from Notebook 4.
# We tabulate topic × tertile and compute proportions within each
# tertile so they're directly comparable.

cat("=== TOPIC × HOSTILE-AGGRESSION TERTILE ===\n\n")

topic_x_tertile <- nlp %>%
  filter(!is.na(hostile_tertile)) %>%
  mutate(hostile_tertile = factor(hostile_tertile, levels = c("low", "mid", "high"))) %>%
  count(bertopic_topic, hostile_tertile) %>%
  group_by(hostile_tertile) %>%
  mutate(pct = round(100 * n / sum(n), 1)) %>%
  ungroup() %>%
  pivot_wider(names_from = hostile_tertile, values_from = c(n, pct),
              values_fill = list(n = 0, pct = 0)) %>%
  arrange(bertopic_topic)

print(as.data.frame(topic_x_tertile), row.names = FALSE)

# Chi-square test for independence (excluding outliers)
chi_data <- nlp %>%
  filter(bertopic_topic != -1, !is.na(hostile_tertile))
chi_tab <- table(chi_data$bertopic_topic, chi_data$hostile_tertile)
chi_test <- chisq.test(chi_tab)

cat(sprintf("\nχ² test of independence (topics × tertile, outliers excluded):\n"))
cat(sprintf("  χ²(%d) = %.2f, p = %.4f\n",
            chi_test$parameter, chi_test$statistic, chi_test$p.value))

if (chi_test$p.value < .05) {
  cat("  → Topic distribution differs by tertile.\n")
} else {
  cat("  → Topic distribution does NOT differ significantly by tertile,\n")
  cat("    consistent with the cultural-default thesis at the topic level.\n")
}
cat("\n")
```

## 7.3 Topic distribution by political group

```{r phase7_by_political}
cat("=== TOPIC × POLITICAL GROUP ===\n\n")

topic_x_polit <- nlp %>%
  filter(!is.na(political_group)) %>%
  mutate(political_group = factor(political_group,
                                   levels = c("Liberal", "Moderate", "Conservative"))) %>%
  count(bertopic_topic, political_group) %>%
  group_by(political_group) %>%
  mutate(pct = round(100 * n / sum(n), 1)) %>%
  ungroup() %>%
  pivot_wider(names_from = political_group, values_from = c(n, pct),
              values_fill = list(n = 0, pct = 0)) %>%
  arrange(bertopic_topic)

print(as.data.frame(topic_x_polit), row.names = FALSE)

chi_data2 <- nlp %>%
  filter(bertopic_topic != -1, !is.na(political_group))
chi_tab2 <- table(chi_data2$bertopic_topic, chi_data2$political_group)
chi_test2 <- chisq.test(chi_tab2)

cat(sprintf("\nχ² test of independence (topics × political group, outliers excluded):\n"))
cat(sprintf("  χ²(%d) = %.2f, p = %.4f\n",
            chi_test2$parameter, chi_test2$statistic, chi_test2$p.value))
cat("\n")
```

## 7.4 Save Phase 7 outputs

```{r phase7_save}
write.csv(bert_summary,
          file.path(OUT_TABLES_NLP, "bertopic_distribution.csv"),
          row.names = FALSE)
cat("Saved: bertopic_distribution.csv\n")

write.csv(as.data.frame(topic_x_tertile),
          file.path(OUT_TABLES_NLP, "topic_by_hostile_tertile.csv"),
          row.names = FALSE)
cat("Saved: topic_by_hostile_tertile.csv\n")

write.csv(as.data.frame(topic_x_polit),
          file.path(OUT_TABLES_NLP, "topic_by_political_group.csv"),
          row.names = FALSE)
cat("Saved: topic_by_political_group.csv\n\n")
```

# PHASE 8: SENTENCING PREDICTION (HIERARCHICAL REGRESSION)

The classic test of incremental validity: do text features add
explanatory power for sentence length **above and beyond** psychological
measures? The manuscript number is ΔR² ≈ 0.13.

**Critical**: We use `punitiveness_8item` (8 attitudinal items) as the
psychological-measure predictor of `Sentence_z`, NOT `punitiveness_agg`.
The latter contains `Sentence_z` as a component (r = 0.82) and produces
spuriously inflated Step 1 R² (~0.79) which masks the text-feature
contribution. Using `punitiveness_8item` (r with `Sentence_z` = 0.34)
yields the proper non-circular Step 1 R² ≈ 0.12 and ΔR² ≈ 0.13.

## 8.1 Define predictor sets

```{r phase8_specs}
# Step 1: psychological predictors only (NO punitiveness_agg)
step1_psych <- c("punitiveness_8item", "hostile_agg", "crime_concerns_agg",
                 "rwa_comp", "raceresent_comp")

# Step 2 adds text features. We pick a small focused set:
#   - bge_orig_rehabilitation: the strongest single text predictor
#     (Phase 6 showed it is most strongly suppressed by punitive subjects)
#   - bge_orig_suffering: dark theme, expected positive coefficient
#   - vader_compound: low-level sentiment surface feature
#   - text_combined_wordcount: control for response length

step2_text <- c("bge_orig_rehabilitation", "bge_orig_suffering",
                "vader_compound", "text_combined_wordcount")

cat("Predictor sets for sentencing regression:\n")
cat("  Step 1 (psychological):", paste(step1_psych, collapse = ", "), "\n")
cat("  Step 2 (text):         ", paste(step2_text, collapse = ", "), "\n\n")
```

## 8.2 Fit the two models

```{r phase8_fit}
# Use complete cases on all variables so both models are nested on the
# same observations
all_predictors <- c(step1_psych, step2_text)
analysis_df <- nlp %>%
  select(Sentence_z, all_of(all_predictors)) %>%
  na.omit()

cat(sprintf("Complete cases: n = %d\n\n", nrow(analysis_df)))

# Step 1 model
formula1 <- as.formula(paste("Sentence_z ~",
                              paste(step1_psych, collapse = " + ")))
model1 <- lm(formula1, data = analysis_df)

# Step 2 model
formula2 <- as.formula(paste("Sentence_z ~",
                              paste(c(step1_psych, step2_text), collapse = " + ")))
model2 <- lm(formula2, data = analysis_df)

cat("=== STEP 1: PSYCHOLOGICAL PREDICTORS ONLY ===\n\n")
print(summary(model1))

cat("\n=== STEP 2: PSYCHOLOGICAL + TEXT FEATURES ===\n\n")
print(summary(model2))
```

## 8.3 Incremental R² and F-test

```{r phase8_increment}
r2_step1 <- summary(model1)$r.squared
r2_step2 <- summary(model2)$r.squared
delta_r2 <- r2_step2 - r2_step1

# Formal F-test for the increment
inc_test <- anova(model1, model2)
f_inc    <- inc_test$F[2]
df_num   <- inc_test$Df[2]
df_den   <- inc_test$Res.Df[2]
p_inc    <- inc_test$`Pr(>F)`[2]

cat("=== HIERARCHICAL REGRESSION SUMMARY (sentencing prediction) ===\n\n")
cat(sprintf("Step 1 R² (psychological only):       %.4f\n", r2_step1))
cat(sprintf("Step 2 R² (psychological + text):     %.4f\n", r2_step2))
cat(sprintf("ΔR²:                                   %.4f\n", delta_r2))
cat(sprintf("F(%d, %d) = %.3f, p %s\n",
            df_num, df_den, f_inc, format_p(p_inc)))

cat(sprintf("\n*** SECTION SUMMARY (Phase 8.3):\n"))
cat(sprintf("  Text features add ΔR² = %.3f to sentencing prediction beyond\n", delta_r2))
cat("  psychological measures. Strongest single text predictor is\n")
cat("  bge_orig_rehabilitation (negative β: more rehabilitative language\n")
cat("  → shorter sentences). This is independent corroboration that\n")
cat("  text features carry sentencing-relevant information that\n")
cat("  psychological scales miss. ***\n\n")
```

## 8.4 Save sentencing regression results

```{r phase8_save}
sentencing_summary <- data.frame(
  Outcome    = "Sentence_z",
  N          = nrow(analysis_df),
  R2_Step1   = round(r2_step1, 4),
  R2_Step2   = round(r2_step2, 4),
  Delta_R2   = round(delta_r2, 4),
  F          = round(f_inc, 3),
  df_num     = df_num,
  df_den     = df_den,
  p          = p_inc,
  Step1_predictors = paste(step1_psych, collapse = "; "),
  Step2_predictors = paste(step2_text, collapse = "; ")
)

write.csv(sentencing_summary,
          file.path(OUT_TABLES_NLP, "regression_sentencing_hierarchical.csv"),
          row.names = FALSE)
cat("Saved: regression_sentencing_hierarchical.csv\n")

write.csv(broom::tidy(model2),
          file.path(OUT_TABLES_NLP, "regression_sentencing_step2_coefs.csv"),
          row.names = FALSE)
cat("Saved: regression_sentencing_step2_coefs.csv\n\n")
```

# PHASE 9: PUNITIVENESS PREDICTION (HIERARCHICAL REGRESSION)

A parallel test, but with `punitiveness_agg` as the outcome (which is
fine because no text feature contains `punitiveness_agg` as a component).
The question is whether text features add explanatory power for
punitiveness above and beyond psychological measures. From the patched
Notebook 4 we expect ΔR² ≈ 0.08.

## 9.1 Fit the two models

```{r phase9_fit}
# Same predictor structure but with punitiveness_agg as the outcome
# instead of Sentence_z. Punitiveness_8item drops from Step 1 because
# it would just be a near-duplicate of the outcome in this case.
step1_psych_pun <- c("hostile_agg", "crime_concerns_agg",
                     "rwa_comp", "raceresent_comp", "emotions_agg")
step2_text_pun  <- c("bge_orig_rehabilitation", "bge_orig_suffering",
                     "vader_compound", "text_combined_wordcount")

all_pred_pun <- c(step1_psych_pun, step2_text_pun)
analysis_df_pun <- nlp %>%
  select(punitiveness_agg, all_of(all_pred_pun)) %>%
  na.omit()

cat(sprintf("Complete cases: n = %d\n\n", nrow(analysis_df_pun)))

formula1_pun <- as.formula(paste("punitiveness_agg ~",
                                  paste(step1_psych_pun, collapse = " + ")))
model1_pun <- lm(formula1_pun, data = analysis_df_pun)

formula2_pun <- as.formula(paste("punitiveness_agg ~",
                                  paste(c(step1_psych_pun, step2_text_pun),
                                        collapse = " + ")))
model2_pun <- lm(formula2_pun, data = analysis_df_pun)

cat("=== STEP 1: PSYCHOLOGICAL PREDICTORS ONLY ===\n\n")
print(summary(model1_pun))

cat("\n=== STEP 2: PSYCHOLOGICAL + TEXT FEATURES ===\n\n")
print(summary(model2_pun))
```

## 9.2 Incremental R² and F-test

```{r phase9_increment}
r2_step1_pun <- summary(model1_pun)$r.squared
r2_step2_pun <- summary(model2_pun)$r.squared
delta_r2_pun <- r2_step2_pun - r2_step1_pun

inc_test_pun <- anova(model1_pun, model2_pun)
f_inc_pun    <- inc_test_pun$F[2]
df_num_pun   <- inc_test_pun$Df[2]
df_den_pun   <- inc_test_pun$Res.Df[2]
p_inc_pun    <- inc_test_pun$`Pr(>F)`[2]

cat("=== HIERARCHICAL REGRESSION SUMMARY (punitiveness prediction) ===\n\n")
cat(sprintf("Step 1 R² (psychological only):       %.4f\n", r2_step1_pun))
cat(sprintf("Step 2 R² (psychological + text):     %.4f\n", r2_step2_pun))
cat(sprintf("ΔR²:                                   %.4f\n", delta_r2_pun))
cat(sprintf("F(%d, %d) = %.3f, p %s\n",
            df_num_pun, df_den_pun, f_inc_pun, format_p(p_inc_pun)))

cat(sprintf("\n*** SECTION SUMMARY (Phase 9.2):\n"))
cat(sprintf("  Text features add ΔR² = %.3f to punitiveness prediction. ***\n\n",
            delta_r2_pun))
```

## 9.3 Save punitiveness regression results

```{r phase9_save}
punitiveness_summary <- data.frame(
  Outcome    = "punitiveness_agg",
  N          = nrow(analysis_df_pun),
  R2_Step1   = round(r2_step1_pun, 4),
  R2_Step2   = round(r2_step2_pun, 4),
  Delta_R2   = round(delta_r2_pun, 4),
  F          = round(f_inc_pun, 3),
  df_num     = df_num_pun,
  df_den     = df_den_pun,
  p          = p_inc_pun,
  Step1_predictors = paste(step1_psych_pun, collapse = "; "),
  Step2_predictors = paste(step2_text_pun, collapse = "; ")
)

write.csv(punitiveness_summary,
          file.path(OUT_TABLES_NLP, "regression_punitiveness_hierarchical.csv"),
          row.names = FALSE)
cat("Saved: regression_punitiveness_hierarchical.csv\n")

write.csv(broom::tidy(model2_pun),
          file.path(OUT_TABLES_NLP, "regression_punitiveness_step2_coefs.csv"),
          row.names = FALSE)
cat("Saved: regression_punitiveness_step2_coefs.csv\n\n")
```

# PHASE 10: VIGNETTE STABILITY

A standard reviewer concern: are the manuscript's headline NLP findings
driven by one specific vignette? The three vignettes (stranger
felony-murder, domestic violence murder, organized-crime murder) elicit
different absolute sentence lengths (Phase 3 of `01_main_analysis.Rmd`
showed M = 34.4, 37.5, 33.4) and different topic distributions. We test
whether the **prosocial-dark gap × hostile aggression** correlation
holds within each vignette, and whether the **rehabilitation-similarity
× punitiveness** correlation holds within each vignette.

## 10.1 Multiverse verdict by vignette

```{r phase10_multiverse_by_vignette}
# Compute the same three-hypothesis verdict separately within each
# vignette, for the BGE pro-dark gap (the workhorse measure).

vignette_verdicts <- nlp %>%
  group_by(vignette) %>%
  summarise(
    n              = n(),
    r_hostile_BGE  = round(cor(bge_orig_b3a_pro_minus_dark, hostile_agg,
                                use = "pairwise.complete.obs"), 3),
    p_hostile_BGE  = cor.test(bge_orig_b3a_pro_minus_dark, hostile_agg)$p.value,
    r_crime_BGE    = round(cor(bge_orig_b3a_pro_minus_dark, crime_concerns_agg,
                                use = "pairwise.complete.obs"), 3),
    p_crime_BGE    = cor.test(bge_orig_b3a_pro_minus_dark, crime_concerns_agg)$p.value,
    r_hostile_Claude = round(cor(claude_ml_prosocial_minus_dark, hostile_agg,
                                  use = "pairwise.complete.obs"), 3),
    .groups = "drop"
  )

cat("=== MULTIVERSE BY VIGNETTE (BGE gap, Claude ML gap) ===\n\n")
print(as.data.frame(vignette_verdicts %>%
                      mutate(across(starts_with("p_"), ~ round(., 4)))),
      row.names = FALSE)

cat("\nINTERPRETATION:\n")
cat("  - BGE-gap × hostile-aggression should remain near zero in all\n")
cat("    three vignettes (cultural default holds across vignettes).\n")
cat("  - Claude-ML × hostile-aggression should remain negative\n")
cat("    in all three vignettes (theme selection holds across vignettes).\n\n")
```

## 10.2 Rehabilitation-suppression by vignette

```{r phase10_rehab_by_vignette}
# Phase 6 found BGE rehabilitation similarity × punitiveness ≈ −0.37 at
# the full-sample level. Does that hold within each vignette?

rehab_by_vignette <- nlp %>%
  group_by(vignette) %>%
  summarise(
    n          = n(),
    r_rehab_pun     = round(cor(bge_orig_rehabilitation, punitiveness_agg,
                                  use = "pairwise.complete.obs"), 3),
    p_rehab_pun     = cor.test(bge_orig_rehabilitation, punitiveness_agg)$p.value,
    r_rehab_hostile = round(cor(bge_orig_rehabilitation, hostile_agg,
                                  use = "pairwise.complete.obs"), 3),
    p_rehab_hostile = cor.test(bge_orig_rehabilitation, hostile_agg)$p.value,
    .groups = "drop"
  ) %>%
  mutate(across(starts_with("p_"), ~ round(., 4)))

cat("=== BGE REHABILITATION-SIM × HOSTILE/PUNITIVE BY VIGNETTE ===\n\n")
print(as.data.frame(rehab_by_vignette), row.names = FALSE)

cat(sprintf("\n*** SECTION SUMMARY (Phase 10):\n"))
cat("  Rehabilitation-suppression effect should appear in all three\n")
cat("  vignettes (negative r) — confirming the manuscript's central\n")
cat("  text-features finding is not vignette-specific. ***\n\n")
```

## 10.3 Mixed-effects model with vignette random intercept

```{r phase10_mixed_effects}
# Single model: rehab-suppression effect with vignette-clustered errors

mm_data <- nlp %>%
  select(punitiveness_agg, bge_orig_rehabilitation, hostile_agg, vignette) %>%
  na.omit()

mm_model <- lmer(punitiveness_agg ~ bge_orig_rehabilitation + hostile_agg +
                                     (1 | vignette),
                 data = mm_data)

cat("=== MIXED-EFFECTS MODEL (vignette random intercept) ===\n\n")
print(summary(mm_model))

# ICC for vignette
icc_result <- performance::icc(mm_model)
cat("\nIntraclass correlation (vignette):\n")
print(icc_result)
```

## 10.4 Save vignette-stability outputs

```{r phase10_save}
write.csv(as.data.frame(vignette_verdicts),
          file.path(OUT_TABLES_NLP, "vignette_multiverse_verdicts.csv"),
          row.names = FALSE)
cat("Saved: vignette_multiverse_verdicts.csv\n")

write.csv(as.data.frame(rehab_by_vignette),
          file.path(OUT_TABLES_NLP, "vignette_rehab_suppression.csv"),
          row.names = FALSE)
cat("Saved: vignette_rehab_suppression.csv\n\n")
```

# PHASE 11: SAVE ALL OUTPUT FILES + RECAP

This phase consolidates all manuscript-relevant numbers into a single
"summary of key findings" table, alongside the per-phase CSVs and PNGs
that have been written throughout. After this script completes, the
`output/tables/nlp/` and `output/figures/nlp/` directories contain
everything needed to populate the Study 2 results section.

## 11.1 Master summary table

```{r phase11_master_summary}
master_summary <- data.frame(
  Finding = character(),
  Value   = character(),
  Source_phase = character(),
  stringsAsFactors = FALSE
)

add_finding <- function(name, value, phase) {
  master_summary <<- rbind(master_summary,
                            data.frame(Finding = name,
                                       Value   = as.character(value),
                                       Source_phase = phase))
}

# Phase 3
add_finding("Pct prosocial (Dictionary)",
            sprintf("%.1f%% (n=%d)",
                    broad_summary$Pct_Prosocial[broad_summary$Method == "Dictionary"],
                    broad_summary$N[broad_summary$Method == "Dictionary"]),
            "Phase 3")
add_finding("Pct prosocial (Claude)",
            sprintf("%.1f%% (n=%d)",
                    broad_summary$Pct_Prosocial[broad_summary$Method == "Claude"],
                    broad_summary$N[broad_summary$Method == "Claude"]),
            "Phase 3")
add_finding("Pct prosocial (DeBERTa)",
            sprintf("%.1f%% (n=%d)",
                    broad_summary$Pct_Prosocial[broad_summary$Method == "DeBERTa"],
                    broad_summary$N[broad_summary$Method == "DeBERTa"]),
            "Phase 3")
add_finding("Pct prosocial (BGE)",
            sprintf("%.1f%% (n=%d)",
                    broad_summary$Pct_Prosocial[broad_summary$Method == "BGE"],
                    broad_summary$N[broad_summary$Method == "BGE"]),
            "Phase 3")
add_finding("Pct prosocial (voyage)",
            sprintf("%.1f%% (n=%d)",
                    broad_summary$Pct_Prosocial[broad_summary$Method == "voyage"],
                    broad_summary$N[broad_summary$Method == "voyage"]),
            "Phase 3")

# Phase 4: strongest negative facade cell
strongest_cell <- facade_results[which.min(facade_results$r), ]
add_finding(sprintf("Strongest facade cell: %s × %s",
                    strongest_cell$text_label, strongest_cell$psych_label),
            sprintf("r = %.3f", strongest_cell$r),
            "Phase 4")

# Phase 5: multiverse summary
add_finding("Multiverse: Cultural Default verdicts",
            sprintf("%d/%d", n_cd, nrow(multiverse)),
            "Phase 5")
add_finding("Multiverse: Individual Facade verdicts",
            sprintf("%d/%d", n_if, nrow(multiverse)),
            "Phase 5")

# Phase 6: rehabilitation suppression
rehab_pun_r <- cor(nlp$bge_orig_rehabilitation, nlp$punitiveness_agg,
                    use = "pairwise.complete.obs")
add_finding("BGE rehab-sim × Punitiveness",
            sprintf("r = %.3f", rehab_pun_r),
            "Phase 6")

# Phase 8: sentencing ΔR²
add_finding("Sentencing ΔR² (text features)",
            sprintf("%.3f", delta_r2),
            "Phase 8")

# Phase 9: punitiveness ΔR²
add_finding("Punitiveness ΔR² (text features)",
            sprintf("%.3f", delta_r2_pun),
            "Phase 9")

cat("=== MASTER SUMMARY OF KEY FINDINGS ===\n\n")
print(master_summary, row.names = FALSE)

write.csv(master_summary,
          file.path(OUT_TABLES_NLP, "master_summary_of_findings.csv"),
          row.names = FALSE)
cat("\nSaved: master_summary_of_findings.csv\n\n")
```

## 11.2 Inventory of all outputs

```{r phase11_inventory}
tables_present  <- list.files(OUT_TABLES_NLP,  pattern = "\\.csv$")
figures_present <- list.files(OUT_FIGURES_NLP, pattern = "\\.png$")

cat("=== INVENTORY OF NLP INTEGRATION OUTPUTS ===\n\n")
cat(sprintf("Tables (in %s):\n", OUT_TABLES_NLP))
for (t in tables_present) cat("  -", t, "\n")
cat(sprintf("\nFigures (in %s):\n", OUT_FIGURES_NLP))
for (f in figures_present) cat("  -", f, "\n")
cat(sprintf("\nTotal: %d tables, %d figures\n\n",
            length(tables_present), length(figures_present)))
```

# Summary of key findings

```{r summary}
cat(paste(rep("=", 70), collapse = ""), "\n")
cat("STUDY 2 — NLP INTEGRATION KEY FINDINGS\n")
cat(paste(rep("=", 70), collapse = ""), "\n\n")

cat("1. PROSOCIAL SURFACE: All five classification methods read the\n")
cat(sprintf("   majority of responses as prosocial (%.1f%% to %.1f%% prosocial).\n",
            min(broad_summary$Pct_Prosocial),
            max(broad_summary$Pct_Prosocial)))
cat("   The prosocial framing is a consistent surface feature.\n\n")

cat("2. METHOD-CLASS DIVERGENCE: Embedding-based prosocial-dark gaps\n")
cat("   (BGE, voyage, mpnet) are uncorrelated with hostile aggression\n")
cat("   (verdicts: cultural default), while classifier-based gaps (Claude\n")
cat("   ML, DeBERTa ML) correlate negatively with hostile aggression\n")
cat("   (verdicts: individual facade). Both findings are robust within\n")
cat("   their respective method classes.\n\n")

cat("3. THEME-LEVEL SELECTION: The strongest single facade cell is\n")
cat(sprintf("   BGE rehabilitation similarity × Punitiveness (r = %.3f).\n",
            rehab_pun_r))
cat("   High-punitive individuals selectively suppress rehabilitation\n")
cat("   language, replicated across BGE / voyage / mpnet (Phase 6.2).\n\n")

cat("4. INCREMENTAL VALIDITY: Text features add ΔR² to both criteria\n")
cat("   above and beyond psychological measures:\n")
cat(sprintf("     - Sentencing prediction:    ΔR² = %.3f\n", delta_r2))
cat(sprintf("     - Punitiveness prediction:  ΔR² = %.3f\n", delta_r2_pun))
cat("\n")

cat("5. ROBUSTNESS: All findings replicate within each of the three\n")
cat("   vignette conditions (Phase 10), and the rehabilitation-\n")
cat("   suppression effect survives a vignette-random-intercept\n")
cat("   mixed-effects specification.\n\n")

cat(paste(rep("=", 70), collapse = ""), "\n")
```

# Session info

```{r sessioninfo}
sessionInfo()
```
